---
title: "Make the prediction grid"
author: "Max Lindmark"
date: "`r format(Sys.time(), '%d %B, %Y')`"
format: 
  html:
    embed-resources: true
    fig-width: 8
    fig-asp: 0.618
knitr: 
  opts_chunk:
    fig.align: center
editor: source
execute: 
  echo: true
  eval: true
  cache: false
---

## Intro
Make an evenly spaced UTM prediction grid with all spatially varying covariates for the diet and the biomass data

```{r lib}
#| message: false

pkgs <- c("tidyverse", "tidylog", "sp", "raster", "devtools", "RCurl", "sdmTMB", "terra", "ncdf4", "chron") 

if(length(setdiff(pkgs, rownames(installed.packages()))) > 0){

    install.packages(setdiff(pkgs, rownames(installed.packages())), dependencies = T)
  
}

invisible(lapply(pkgs, library, character.only = T))

# Source code for map plots
source_url("https://raw.githubusercontent.com/maxlindmark/cod-interactions/main/R/functions/map-plot.R")

# Packages not on CRAN
# devtools::install_github("seananderson/ggsidekick") # not on CRAN 
library(ggsidekick)
theme_set(theme_sleek())

# Set path
home <- here::here()
```

Read data and depth-raster

```{r data}
# Read data
d <- readr::read_csv("https://raw.githubusercontent.com/maxlindmark/spatial-metabolic-index/main/data/clean/catch_clean.csv") |> 
  rename(X = x, Y = y)
```

## Make the grid with depth
First make a grid for the biomass data, then subset that based on the extend of the stomach data

```{r make pred grid}
x <- d$X
y <- d$Y
z <- chull(x, y)

coords <- cbind(x[z], y[z])

coords <- rbind(coords, coords[1, ])

plot(coords[, 1] ~ coords[, 2]) # plot data

sp_poly <- sp::SpatialPolygons(
  list(sp::Polygons(list(sp::Polygon(coords)), ID = 1))
  )

sp_poly_df <- sp::SpatialPolygonsDataFrame(sp_poly,
                                           data = data.frame(ID = 1)
                                           )
cell_width <- 3

pred_grid <- expand.grid(
  X = seq(min(d$X), max(d$X), cell_width),
  Y = seq(min(d$Y), max(d$Y), cell_width),
  year = c(1993:2021)
  )

ggplot(pred_grid |> filter(year == 2019), aes(X, Y)) +
  geom_point(size = 0.1) +
  theme_void() +
  coord_sf()

sp::coordinates(pred_grid) <- c("X", "Y")

inside <- !is.na(sp::over(pred_grid, as(sp_poly_df, "SpatialPolygons")))

pred_grid <- pred_grid[inside, ]

pred_grid <- as.data.frame(pred_grid)

ggplot(data = filter(pred_grid, year == 1999), aes(X*1000, Y*1000)) + 
  geom_point(size = 0.001, alpha = 0.5) +
  NULL

plot_map +
  geom_point(data = filter(pred_grid, year == 1999), aes(X*1000, Y*1000), size = 0.001, alpha = 0.5) +
  NULL

# Add lat and lon
# Need to go from UTM to lat long for this one...
# https://stackoverflow.com/questions/30018098/how-to-convert-utm-coordinates-to-lat-and-long-in-r
xy <- as.matrix(pred_grid |> dplyr::select(X, Y) |> mutate(X = X*1000, Y = Y*1000))
v <- vect(xy, crs="+proj=utm +zone=33 +datum=WGS84  +units=m")
y <- project(v, "+proj=longlat +datum=WGS84")
lonlat <- geom(y)[, c("x", "y")]

pred_grid$lon <- lonlat[, 1]
pred_grid$lat <- lonlat[, 2]

ggplot(filter(pred_grid, year == 1999), aes(lon, lat)) + geom_point()

# Add depth now to remove islands and remaining land
# https://gis.stackexchange.com/questions/411261/read-multiple-layers-raster-from-ncdf-file-using-terra-package
# https://emodnet.ec.europa.eu/geoviewer/
dep_raster <- terra::rast(paste0(home, "/data/Mean depth natural colour (with land).nc"))
class(dep_raster)
crs(dep_raster, proj = TRUE)

plot(dep_raster)

pred_grid$depth <- terra::extract(dep_raster, pred_grid |> dplyr::select(lon, lat))$elevation

ggplot(pred_grid, aes(lon, lat, color = depth*-1)) + 
  geom_point()

pred_grid$depth <- pred_grid$depth*-1

pred_grid <- pred_grid |> drop_na(depth)

pred_grid |> 
  filter(year == 1999) |> 
  drop_na(depth) |> 
  #mutate(water = ifelse(depth < 0.00000001, "N", "Y")) |> 
  ggplot(aes(X*1000, Y*1000, fill = depth)) + 
  geom_raster() +
  NULL

plot_map + 
  geom_point(data = pred_grid, aes(X*1000, Y*1000), size = 0.001) + 
  geom_sf()

plot_map + 
  geom_raster(data = filter(pred_grid, year == 1999), aes(X*1000, Y*1000, fill = depth), size = 0.001) + 
  geom_sf()
```

### Substrate
First add lat and lon based on X and Y (utm)

```{r substrate}
substrate <- terra::rast(paste0(home, "/data/substrate_tif/BALANCE_SEABED_SEDIMENT.tif"))

newcrs <- "+proj=longlat +datum=WGS84"

substrate_longlat = terra::project(substrate, newcrs)
 
plot(substrate_longlat)

# Now extract the values from the saduria raster to pred_grid
pred_grid$substrate <- terra::extract(substrate_longlat, pred_grid |> dplyr::select(lon, lat))$BALANCE_SEABED_SEDIMENT

unique(pred_grid$substrate)

factor(sort(unique(round(pred_grid$substrate))))
 
pred_grid$substrate <- round(pred_grid$substrate)

pred_grid <- pred_grid %>% mutate(substrate = ifelse(substrate == 1, "bedrock", substrate),
                                  substrate = ifelse(substrate == 2, "hard-bottom complex", substrate),
                                  substrate = ifelse(substrate == 3, "sand", substrate),
                                  substrate = ifelse(substrate == 4, "hard clay", substrate),
                                  substrate = ifelse(substrate == 5, "mud", substrate))

# I. Bedrock.
# II. Hard bottom complex, includes patchy hard surfaces and coarse sand (sometimes also clay) to boulders.
# III. Sand including fine to coarse sand (with gravel exposures).
# IV. Hard clay sometimes/often/possibly exposed or covered with a thin layer of sand/gravel.
# V. Mud including gyttja-clay to gyttja-silt.

# Plot
ggplot(pred_grid, aes(X, Y, fill = substrate)) +
  geom_raster() +
  coord_sf()
```

Add month_year variable for matching with raster layers

```{r merge quarters}
pred_grid_1 <- pred_grid |> mutate(quarter = 1)
pred_grid_4 <- pred_grid |> mutate(quarter = 4)

dat <- bind_rows(pred_grid_1, pred_grid_4) |> 
  mutate(month = ifelse(quarter == 1, 2, 11), # most common months
         month_year = paste(month, year, sep = "_"))
```

## Oxygen

```{r oxygen}
# Downloaded from here: https://data.marine.copernicus.eu/products
# Extract raster points: https://gisday.wordpress.com/2014/03/24/extract-raster-values-from-points-using-r/comment-page-1/
# https://rpubs.com/boyerag/297592
# https://pjbartlein.github.io/REarthSysSci/netCDF.html#get-a-variable
# Open the netCDF file
ncin <- nc_open(paste0(home, "/data/NEMO/cmems_mod_bal_bgc_my_P1M-m_1691065496556.nc"))

print(ncin)

# Get longitude and latitude
lon <- ncvar_get(ncin,"lon")
nlon <- dim(lon)
head(lon)

lat <- ncvar_get(ncin,"lat")
nlat <- dim(lat)
head(lat)

# Get time
time <- ncvar_get(ncin,"time")
time

tunits <- ncatt_get(ncin,"time","units")
nt <- dim(time)
nt
tunits

# Get oxygen
dname <- "o2b"

oxy_array <- ncvar_get(ncin,dname)
dlname <- ncatt_get(ncin,dname,"long_name")
dunits <- ncatt_get(ncin,dname,"units")
fillvalue <- ncatt_get(ncin,dname,"_FillValue")
dim(oxy_array)

# Get global attributes
title <- ncatt_get(ncin,0,"title")
institution <- ncatt_get(ncin,0,"institution")
datasource <- ncatt_get(ncin,0,"source")
references <- ncatt_get(ncin,0,"references")
history <- ncatt_get(ncin,0,"history")
Conventions <- ncatt_get(ncin,0,"Conventions")

# Convert time: split the time units string into fields
tustr <- strsplit(tunits$value, " ")
tdstr <- strsplit(unlist(tustr)[3], "-")
tmonth <- as.integer(unlist(tdstr)[2])
tday <- as.integer(unlist(tdstr)[3])
tyear <- as.integer(unlist(tdstr)[1])

# Here I deviate from the guide a little bit. Save this info:
dates <- chron(time, origin = c(tmonth, tday, tyear))

# Crop the date variable
months <- as.numeric(substr(dates, 2, 3))
years <- as.numeric(substr(dates, 8, 9))
years <- ifelse(years > 90, 1900 + years, 2000 + years)

# Replace netCDF fill values with NA's
oxy_array[oxy_array == fillvalue$value] <- NA

dim(oxy_array)
str(dim(oxy_array))
# The third slot is the date index

# Loop through all "dates", put into a list 
dlist <- list()

for(i in 1:length(months)) {
  
  oxy_sub <- oxy_array[, , i]
    
  dlist[[i]] <- oxy_sub
  
}

# Name the list
names(dlist) <- paste(months, years, sep = "_")
str(dlist)

# Create data holding object
oxy_data_list <- list()

# Loop through each month_year and extract raster values for the cpue data points
for(i in unique(dat$month_year)) {
  
  # Set plot limits
  ymin = 54; ymax = 58; xmin = 12; xmax = 22

  # Subset a month-year combination
  oxy_slice <- dlist[[i]]
  
  # Create raster for that year (i)
  r <- raster(t(oxy_slice), xmn = min(lon), xmx = max(lon), ymn = min(lat), ymx = max(lat),
              crs = CRS("+proj=longlat +ellps=WGS84 +datum=WGS84 +no_defs+ towgs84=0,0,0"))
  
  # Flip...
  r <- flip(r, direction = 'y')
  
  plot(r, main = paste(i), ylim = c(ymin, ymax), xlim = c(xmin, xmax))

  # Filter the same year (i) in the cpue data and select only coordinates
  d_slice <- dat %>% filter(month_year == i) %>% dplyr::select(lon, lat)
  
  # Make into a SpatialPoints object
  data_sp <- SpatialPoints(d_slice)
  
  # Extract raster value (oxygen)
  rasValue <- raster::extract(r, data_sp)
  
  # Now we want to plot the results of the raster extractions by plotting the cpue data points over a raster and saving it for each year.
  # Make the SpatialPoints object into a raster again (for plot)
  df <- as.data.frame(data_sp)
  
  # Add in the raster value in the df holding the coordinates for the cpue data
  d_slice$oxy <- rasValue
  
  # Add in which year
  d_slice$month_year <- i

  # Now the unit of oxygen is mmol/m3. I want it to be ml/L. The original model is in unit ml/L
  # and it's been converted by the data host. Since it was converted without accounting for
  # pressure or temperature, I can simply use the following conversion factor:
  # 1 ml/l = 103/22.391 = 44.661 μmol/l -> 1 ml/l = 0.044661 mmol/l = 44.661 mmol/m^3 -> 0.0223909 ml/l = 1mmol/m^3
  # https://ocean.ices.dk/tools/unitconversion.aspx

  d_slice$oxy <- d_slice$oxy * 0.0223909
    
  # Add each years' data in the list
  oxy_data_list[[i]] <- d_slice

}

# Now create a data frame from the list of all annual values
big_dat_oxy <- dplyr::bind_rows(oxy_data_list)
```

#### Temperature

```{r temperature}
# Open the netCDF file
ncin <- nc_open(paste0(home, "/data/NEMO/cmems_mod_bal_phy_my_P1M-m_1691065935335.nc"))
                                        
print(ncin)

# Get longitude and latitude
lon <- ncvar_get(ncin,"lon")
nlon <- dim(lon)
head(lon)

lat <- ncvar_get(ncin,"lat")
nlat <- dim(lat)
head(lat)

# Get time
time <- ncvar_get(ncin,"time")
time

tunits <- ncatt_get(ncin,"time","units")
nt <- dim(time)
nt
tunits

# Get temperature
dname <- "bottomT"

temp_array <- ncvar_get(ncin,dname)
dlname <- ncatt_get(ncin,dname,"long_name")
dunits <- ncatt_get(ncin,dname,"units")
fillvalue <- ncatt_get(ncin,dname,"_FillValue")
dim(temp_array)

# Get global attributes
title <- ncatt_get(ncin,0,"title")
institution <- ncatt_get(ncin,0,"institution")
datasource <- ncatt_get(ncin,0,"source")
references <- ncatt_get(ncin,0,"references")
history <- ncatt_get(ncin,0,"history")
Conventions <- ncatt_get(ncin,0,"Conventions")

# Convert time: split the time units string into fields
tustr <- strsplit(tunits$value, " ")
tdstr <- strsplit(unlist(tustr)[3], "-")
tmonth <- as.integer(unlist(tdstr)[2])
tday <- as.integer(unlist(tdstr)[3])
tyear <- as.integer(unlist(tdstr)[1])

# Here I deviate from the guide a little bit. Save this info:
dates <- chron(time, origin = c(tmonth, tday, tyear))

# Crop the date variable
months <- as.numeric(substr(dates, 2, 3))
years <- as.numeric(substr(dates, 8, 9))
years <- ifelse(years > 90, 1900 + years, 2000 + years)

# Replace netCDF fill values with NA's
temp_array[temp_array == fillvalue$value] <- NA

# Loop through all "dates", put into a list 
dlist <- list()

for(i in 1:length(months)) {
  
  temp_sub <- temp_array[, , i]
  
  dlist[[i]] <- temp_sub
  
}

# Name the list
names(dlist) <- paste(months, years, sep = "_")
str(dlist)

# Create data holding object
temp_data_list <- list()

# Loop through each month_year and extract raster values for the cpue data points
for(i in unique(dat$month_year)) { # We can use q1 as looping index, doesn't matter!
  
  # Set plot limits
  ymin = 54; ymax = 58; xmin = 12; xmax = 22
  
  # Subset a month-year combination
  temp_slice <- dlist[[i]]
  
  # Create raster for that year (i)
  r <- raster(t(temp_slice), xmn = min(lon), xmx = max(lon), ymn = min(lat), ymx = max(lat),
              crs = CRS("+proj=longlat +ellps=WGS84 +datum=WGS84 +no_defs+ towgs84=0,0,0"))
  
  # Flip...
  r <- flip(r, direction = 'y')
  
  plot(r, main = paste(i), ylim = c(ymin, ymax), xlim = c(xmin, xmax))
  
  # Filter the same year (i) in the cpue data and select only coordinates
  d_slice <- dat %>% filter(month_year == i) %>% dplyr::select(lon, lat)
  
  # Make into a SpatialPoints object
  data_sp <- SpatialPoints(d_slice)
  
  # Extract raster value (oxygen)
  rasValue <- raster::extract(r, data_sp)
  
  # Now we want to plot the results of the raster extractions by plotting the cpue data points over a raster and saving it for each year.
  # Make the SpatialPoints object into a raster again (for plot)
  df <- as.data.frame(data_sp)
  
  # Add in the raster value in the df holding the coordinates for the cpue data
  d_slice$temp <- rasValue
  
  # Add in which year
  d_slice$month_year <- i
  
  # Add each years' data in the list
  temp_data_list[[i]] <- d_slice
  
}

# Now create a data frame from the list of all annual values
big_dat_temp <- dplyr::bind_rows(temp_data_list)
```

#### Salinity

```{r salinity}
# Open the netCDF file
ncin <- nc_open(paste0(home, "/data/NEMO/cmems_mod_bal_phy_my_P1M-m_1691065974147.nc"))

print(ncin)

# Get longitude and latitude
lon <- ncvar_get(ncin,"lon")
nlon <- dim(lon)
head(lon)

lat <- ncvar_get(ncin,"lat")
nlat <- dim(lat)
head(lat)

# Get time
time <- ncvar_get(ncin,"time")
time

tunits <- ncatt_get(ncin,"time","units")
nt <- dim(time)
nt
tunits

# Get Salinity
dname <- "sob"

sal_array <- ncvar_get(ncin,dname)
dlname <- ncatt_get(ncin,dname,"long_name")
dunits <- ncatt_get(ncin,dname,"units")
fillvalue <- ncatt_get(ncin,dname,"_FillValue")
dim(sal_array)

# Get global attributes
title <- ncatt_get(ncin,0,"title")
institution <- ncatt_get(ncin,0,"institution")
datasource <- ncatt_get(ncin,0,"source")
references <- ncatt_get(ncin,0,"references")
history <- ncatt_get(ncin,0,"history")
Conventions <- ncatt_get(ncin,0,"Conventions")

# Convert time: split the time units string into fields
tustr <- strsplit(tunits$value, " ")
tdstr <- strsplit(unlist(tustr)[3], "-")
tmonth <- as.integer(unlist(tdstr)[2])
tday <- as.integer(unlist(tdstr)[3])
tyear <- as.integer(unlist(tdstr)[1])

# Here I deviate from the guide a little bit. Save this info:
dates <- chron(time, origin = c(tmonth, tday, tyear))

# Crop the date variable
months <- as.numeric(substr(dates, 2, 3))
years <- as.numeric(substr(dates, 8, 9))
years <- ifelse(years > 90, 1900 + years, 2000 + years)

# Replace netCDF fill values with NA's
sal_array[sal_array == fillvalue$value] <- NA

# Loop through all "dates", put into a list 
dlist <- list()

for(i in 1:length(months)) {
  
  sal_sub <- sal_array[, , i]
  
  dlist[[i]] <- sal_sub
  
}

# Name the list
names(dlist) <- paste(months, years, sep = "_")
str(dlist)

# Create data holding object
sal_data_list <- list()

# Loop through each month_year and extract raster values for the cpue data points
for(i in unique(dat$month_year)) { # We can use q1 as looping index, doesn't matter!
  
  # Set plot limits
  ymin = 54; ymax = 58; xmin = 12; xmax = 22
  
  # Subset a month-year combination
  sal_slice <- dlist[[i]]
  
  # Create raster for that year (i)
  r <- raster(t(sal_slice), xmn = min(lon), xmx = max(lon), ymn = min(lat), ymx = max(lat),
              crs = CRS("+proj=longlat +ellps=WGS84 +datum=WGS84 +no_defs+ towgs84=0,0,0"))
  
  # Flip...
  r <- flip(r, direction = 'y')
  
  plot(r, main = paste(i), ylim = c(ymin, ymax), xlim = c(xmin, xmax))
  
  # Filter the same year (i) in the cpue data and select only coordinates
  d_slice <- dat %>% filter(month_year == i) %>% dplyr::select(lon, lat)
  
  # Make into a SpatialPoints object
  data_sp <- SpatialPoints(d_slice)
  
  # Extract raster value (oxygen)
  rasValue <- raster::extract(r, data_sp)
  
  # Now we want to plot the results of the raster extractions by plotting the cpue data points over a raster and saving it for each year.
  # Make the SpatialPoints object into a raster again (for plot)
  df <- as.data.frame(data_sp)
  
  # Add in the raster value in the df holding the coordinates for the cpue data
  d_slice$sal <- rasValue
  
  # Add in which year
  d_slice$month_year <- i
  
  # Add each years' data in the list
  sal_data_list[[i]] <- d_slice
  
}

# Now create a data frame from the list of all annual values
big_dat_sal <- dplyr::bind_rows(sal_data_list)
```

```{r merge oxygen temp and salinity data with fish data}
env_dat <- left_join(big_dat_oxy, big_dat_temp, by = c("month_year", "lon", "lat")) |> 
  left_join(big_dat_sal, by = c("month_year", "lon", "lat"))

# Now join these data with the full_dat
dat_full <- left_join(dat, env_dat, by = c("month_year", "lon", "lat"))

ggplot(dat_full |> filter(quarter == 1), aes(X, Y, color = oxy)) + 
  geom_point() + 
  facet_wrap(~year)
```

## Add ICES areas

```{r ices areas}
# https://stackoverflow.com/questions/34272309/extract-shapefile-value-to-point-with-r
# https://gis.ices.dk/sf/
shape <- shapefile(paste0(home, "/data/ICES_StatRec_mapto_ICES_Areas/StatRec_map_Areas_Full_20170124.shp"))
head(shape)

pts <- SpatialPoints(cbind(dat_full$lon, dat_full$lat), 
                     proj4string = CRS(proj4string(shape)))

dat_full$subdiv <- over(pts, shape)$Area_27

# Rename subdivisions to the more common names and do some more filtering (by sub div and area)
sort(unique(dat_full$subdiv))

dat_full <- dat_full |> 
  mutate(sub_div = factor(subdiv),
         sub_div = fct_recode(subdiv,
                              "24" = "3.d.24",
                              "25" = "3.d.25",
                              "26" = "3.d.26",
                              "27" = "3.d.27",
                              "28" = "3.d.28.1",
                              "28" = "3.d.28.2",
                              "29" = "3.d.29"),
         sub_div = as.character(sub_div)) |> 
  filter(sub_div %in% c("24", "25", "26", "27", "28", 2)) |> 
  filter(lat > 54 & lat < 59 & lon < 22)

# Add ICES rectangles
dat_full$ices_rect <- mapplots::ices.rect2(lon = dat_full$lon, lat = dat_full$lat)

plot_map +
  geom_raster(data = filter(dat_full, year == 1999), aes(X*1000, Y*1000, fill = oxy)) +
  facet_wrap(~sub_div)

dat_full <- dat_full |> dplyr::select(-subdiv)
```

## Save

```{r save}
# Remove variables and save
pred_grid_93_06 <- dat_full |> filter(year < 2007)
pred_grid_07_19 <- dat_full |> filter(year > 2006)

write_csv(pred_grid_93_06, file = paste0(home, "/data/clean/pred_grid_(1_2).csv"))
write_csv(pred_grid_07_19, file = paste0(home, "/data/clean/pred_grid_(2_2).csv"))
```

```{r}
knitr::knit_exit()
```

## Compare the old NEMO-NORDIC with the new one
#### Oxygen

```{r}
# Downloaded from here: https://resources.marine.copernicus.eu/?option=com_csw&view=details&product_id=BALTICSEA_REANALYSIS_BIO_003_012
# Extract raster points: https://gisday.wordpress.com/2014/03/24/extract-raster-values-from-points-using-r/comment-page-1/
# https://rpubs.com/boyerag/297592
# https://pjbartlein.github.io/REarthSysSci/netCDF.html#get-a-variable
# Open the netCDF file
ncin <- nc_open(paste0(home, "/data/NEMO/old/dataset-reanalysis-scobi-monthlymeans_1664182224542.nc"))

print(ncin)

# Get longitude and latitude
lon <- ncvar_get(ncin,"longitude")
nlon <- dim(lon)
head(lon)

lat <- ncvar_get(ncin,"latitude")
nlat <- dim(lat)
head(lat)

# Get time
time <- ncvar_get(ncin,"time")
time

tunits <- ncatt_get(ncin,"time","units")
nt <- dim(time)
nt
tunits

# Get oxygen
dname <- "o2b"

oxy_array <- ncvar_get(ncin,dname)
dlname <- ncatt_get(ncin,dname,"long_name")
dunits <- ncatt_get(ncin,dname,"units")
fillvalue <- ncatt_get(ncin,dname,"_FillValue")
dim(oxy_array)

# Get global attributes
title <- ncatt_get(ncin,0,"title")
institution <- ncatt_get(ncin,0,"institution")
datasource <- ncatt_get(ncin,0,"source")
references <- ncatt_get(ncin,0,"references")
history <- ncatt_get(ncin,0,"history")
Conventions <- ncatt_get(ncin,0,"Conventions")

# Convert time: split the time units string into fields
tustr <- strsplit(tunits$value, " ")
tdstr <- strsplit(unlist(tustr)[3], "-")
tmonth <- as.integer(unlist(tdstr)[2])
tday <- as.integer(unlist(tdstr)[3])
tyear <- as.integer(unlist(tdstr)[1])

# Here I deviate from the guide a little bit. Save this info:
dates <- chron(time, origin = c(tmonth, tday, tyear))

# Crop the date variable
months <- as.numeric(substr(dates, 2, 3))
years <- as.numeric(substr(dates, 8, 9))
years <- ifelse(years > 90, 1900 + years, 2000 + years)

# Replace netCDF fill values with NA's
oxy_array[oxy_array == fillvalue$value] <- NA

dim(oxy_array)
str(dim(oxy_array))
# The third slot is the date index

# Loop through all "dates", put into a list 
dlist <- list()

for(i in 1:length(months)) {
  
  oxy_sub <- oxy_array[, , i]
    
  dlist[[i]] <- oxy_sub
  
}

# Name the list
names(dlist) <- paste(months, years, sep = "_")
str(dlist)

# Create data holding object
oxy_data_list <- list()

dat_old <- dat |> filter(!month_year %in% c("11_2020", "2_2021", "11_2021"))

# Loop through each month_year and extract raster values for the cpue data points
for(i in unique(dat_old$month_year)) { # We can use q1 as looping index, doesn't matter!
  
  # Set plot limits
  ymin = 54; ymax = 58; xmin = 12; xmax = 22

  # Subset a month-year combination
  oxy_slice <- dlist[[i]]
  
  # Create raster for that year (i)
  r <- raster(t(oxy_slice), xmn = min(lon), xmx = max(lon), ymn = min(lat), ymx = max(lat),
              crs = CRS("+proj=longlat +ellps=WGS84 +datum=WGS84 +no_defs+ towgs84=0,0,0"))
  
  # Flip...
  r <- flip(r, direction = 'y')
  
  plot(r, main = paste(i), ylim = c(ymin, ymax), xlim = c(xmin, xmax))

  # Filter the same year (i) in the cpue data and select only coordinates
  d_slice <- dat_old %>% filter(month_year == i) %>% dplyr::select(lon, lat)
  
  # Make into a SpatialPoints object
  data_sp <- SpatialPoints(d_slice)
  
  # Extract raster value (oxygen)
  rasValue <- raster::extract(r, data_sp)
  
  # Now we want to plot the results of the raster extractions by plotting the cpue data points over a raster and saving it for each year.
  # Make the SpatialPoints object into a raster again (for plot)
  df <- as.data.frame(data_sp)
  
  # Add in the raster value in the df holding the coordinates for the cpue data
  d_slice$oxy <- rasValue
  
  # Add in which year
  d_slice$month_year <- i

  # Now the unit of oxygen is mmol/m3. I want it to be ml/L. The original model is in unit ml/L
  # and it's been converted by the data host. Since it was converted without accounting for
  # pressure or temperature, I can simply use the following conversion factor:
  # 1 ml/l = 103/22.391 = 44.661 μmol/l -> 1 ml/l = 0.044661 mmol/l = 44.661 mmol/m^3 -> 0.0223909 ml/l = 1mmol/m^3
  # https://ocean.ices.dk/tools/unitconversion.aspx

  d_slice$oxy <- d_slice$oxy * 0.0223909
    
  # Add each years' data in the list
  oxy_data_list[[i]] <- d_slice

}

# Now create a data frame from the list of all annual values
big_dat_oxy <- dplyr::bind_rows(oxy_data_list)

# Merge oxygen temp and salinity data with fish data}
big_dat_oxy <- big_dat_oxy |> rename(oxy_old = oxy)

dat_full2 <- left_join(dat_full, big_dat_oxy, by = c("month_year", "lon", "lat"))

ggplot(dat_full2 |> filter(year == 2000), aes(oxy, oxy_old)) + 
  geom_point() + 
  geom_abline(color = "red")

min(dat_full2$oxy_old, na.rm = TRUE)
min(dat_full2$oxy, na.rm = TRUE)
dat_full2 |> arrange(oxy)
hist(dat_full2$oxy_old)
hist(dat_full2$oxy)

ggplot(dat_full2, aes(oxy, oxy_old)) + 
  geom_point(alpha = 0.1) + 
  facet_wrap(~year)

cordf <- dat_full2 |> drop_na(oxy, oxy_old)

cor(cordf$oxy_old, cordf$oxy)

dat_full2 |> 
  mutate(diff = oxy - oxy_old) |> 
  ggplot(aes(X, Y, fill = diff)) + 
  facet_wrap(~year) + 
  geom_raster() + 
  scale_fill_gradient2()


dat_full2 |> 
  mutate(diff = oxy - oxy_old) |> 
  mutate(diff = ifelse(diff < 5, diff, NA)) |> 
  ggplot(aes(X, Y, fill = diff)) + 
  facet_wrap(~year) + 
  geom_raster() + 
  scale_fill_gradient2()

dat_full2 |> 
  mutate(diff = oxy - oxy_old) |> 
  mutate(diff = ifelse(diff < 5, diff, NA)) |> 
  ggplot(aes(depth, diff)) + 
  geom_point(alpha = 0.1) +
  labs(y = "oxy - oxy_old")
```

#### Temperature

```{r}
# Open the netCDF file
ncin <- nc_open(paste0(home, "/data/NEMO/old/dataset-reanalysis-nemo-monthlymeans_1664183191233.nc"))
                                        
print(ncin)

# Get longitude and latitude
lon <- ncvar_get(ncin,"longitude")
nlon <- dim(lon)
head(lon)

lat <- ncvar_get(ncin,"latitude")
nlat <- dim(lat)
head(lat)

# Get time
time <- ncvar_get(ncin,"time")
time

tunits <- ncatt_get(ncin,"time","units")
nt <- dim(time)
nt
tunits

# Get temperature
dname <- "bottomT"

temp_array <- ncvar_get(ncin,dname)
dlname <- ncatt_get(ncin,dname,"long_name")
dunits <- ncatt_get(ncin,dname,"units")
fillvalue <- ncatt_get(ncin,dname,"_FillValue")
dim(temp_array)

# Get global attributes
title <- ncatt_get(ncin,0,"title")
institution <- ncatt_get(ncin,0,"institution")
datasource <- ncatt_get(ncin,0,"source")
references <- ncatt_get(ncin,0,"references")
history <- ncatt_get(ncin,0,"history")
Conventions <- ncatt_get(ncin,0,"Conventions")

# Convert time: split the time units string into fields
tustr <- strsplit(tunits$value, " ")
tdstr <- strsplit(unlist(tustr)[3], "-")
tmonth <- as.integer(unlist(tdstr)[2])
tday <- as.integer(unlist(tdstr)[3])
tyear <- as.integer(unlist(tdstr)[1])

# Here I deviate from the guide a little bit. Save this info:
dates <- chron(time, origin = c(tmonth, tday, tyear))

# Crop the date variable
months <- as.numeric(substr(dates, 2, 3))
years <- as.numeric(substr(dates, 8, 9))
years <- ifelse(years > 90, 1900 + years, 2000 + years)

# Replace netCDF fill values with NA's
temp_array[temp_array == fillvalue$value] <- NA

# Loop through all "dates", put into a list 
dlist <- list()

for(i in 1:length(months)) {
  
  temp_sub <- temp_array[, , i]
  
  dlist[[i]] <- temp_sub
  
}

# Name the list
names(dlist) <- paste(months, years, sep = "_")
str(dlist)

# Create data holding object
temp_data_list <- list()

dat_old <- dat |> filter(!month_year %in% c("11_2020", "2_2021", "11_2021"))

# Loop through each month_year and extract raster values for the cpue data points
for(i in unique(dat_old$month_year)) { # We can use q1 as looping index, doesn't matter!
  
  # Set plot limits
  ymin = 54; ymax = 58; xmin = 12; xmax = 22
  
  # Subset a month-year combination
  temp_slice <- dlist[[i]]
  
  # Create raster for that year (i)
  r <- raster(t(temp_slice), xmn = min(lon), xmx = max(lon), ymn = min(lat), ymx = max(lat),
              crs = CRS("+proj=longlat +ellps=WGS84 +datum=WGS84 +no_defs+ towgs84=0,0,0"))
  
  # Flip...
  r <- flip(r, direction = 'y')
  
  plot(r, main = paste(i), ylim = c(ymin, ymax), xlim = c(xmin, xmax))
  
  # Filter the same year (i) in the cpue data and select only coordinates
  d_slice <- dat_old %>% filter(month_year == i) %>% dplyr::select(lon, lat)
  
  # Make into a SpatialPoints object
  data_sp <- SpatialPoints(d_slice)
  
  # Extract raster value (oxygen)
  rasValue <- raster::extract(r, data_sp)
  
  # Now we want to plot the results of the raster extractions by plotting the cpue data points over a raster and saving it for each year.
  # Make the SpatialPoints object into a raster again (for plot)
  df <- as.data.frame(data_sp)
  
  # Add in the raster value in the df holding the coordinates for the cpue data
  d_slice$temp <- rasValue
  
  # Add in which year
  d_slice$month_year <- i
  
  # Add each years' data in the list
  temp_data_list[[i]] <- d_slice
  
}

# Now create a data frame from the list of all annual values
big_dat_temp <- dplyr::bind_rows(temp_data_list)

# Merge oxygen temp and salinity data with fish data}
big_dat_temp <- big_dat_temp |> rename(temp_old = temp)

dat_full2 <- left_join(dat_full, big_dat_temp, by = c("month_year", "lon", "lat"))

ggplot(dat_full2 |> filter(year == 2000), aes(temp, temp_old)) + 
  geom_point() + 
  geom_abline(color = "red")

ggplot(dat_full2, aes(temp, temp_old)) + 
  geom_point() + 
  facet_wrap(~year)
```

