---
title: "Fit random forest models to spatial biomass trends and velocities"
author: "Max Lindmark"
date: today
date-format: iso
toc: true
format: 
  html:
    page-layout: full
    embed-resources: true
knitr: 
  opts_chunk:
    fig.align: center
    out-width: 100%
editor: source
---

## Intro

Fit random forest models to biomass trends and velocities from sdms random fields. We'll use a `tidymodels` pipeline, consisting of `rsample` for sample splitting (e.g. train/test or cross-validation), `recipes` for pre-processing, `parsnip` for specifying the model and `yardstick` for evaluating the model.

## Load packages & source functions

```{r load libraries}
#| message: false
#| warning: false

library(tidyverse)
library(tidylog)
library(tidymodels)
library(RCurl)
library(devtools)
library(vip)
library(viridis)
library(RColorBrewer)
library(ranger)
library(patchwork)
library(DALEXtra)
library(modelr)
library(egg)

# Source code for map plots
devtools::source_url("https://raw.githubusercontent.com/maxlindmark/spatial-metabolic-index/main/R/functions/map-plot.R")

# Set path
home <- here::here()

# Set seed
set.seed(42)
```

## Load data

```{r}
#| message: false
#| warning: false

d <- read_csv(paste0(home, "/output/rf_dat_long.csv")) |> 
  filter(cumulative == "Y") |> 
  filter(!group == "dab_all") |> 
  dplyr::select(biom_trend, temp_slope, mean_temp, oxy_slope, mean_oxy, mean_sal, sal_slope, group, mean_biom) |> 
  mutate(mean_biom = (mean_biom - mean(mean_biom)) / sd(mean_biom),
         biom_trend = (biom_trend - mean(biom_trend)) / sd(biom_trend),
         .by = group)
```

```{r}
# Select numeric predictors only
cor_dat <- trend |> 
  select(-biom_trend, -group)

# Correlation matrix
cor_mat <- cor(cor_dat, use = "pairwise.complete.obs")

# Convert to long format
cor_long <- cor_mat |>
  as.data.frame() |>
  rownames_to_column("Var1") |>
  pivot_longer(-Var1, names_to = "Var2", values_to = "Correlation")

# Plot with values
ggplot(cor_long, aes(Var1, Var2, fill = Correlation)) +
  geom_tile() +
  geom_text(aes(label = sprintf("%.2f", Correlation)),
            size = 3) +
  scale_fill_gradient2(low = "steelblue",
                       mid = "white",
                       high = "tomato3",
                       midpoint = 0,
                       limits = c(-1, 1)) +
  coord_equal() +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  labs(fill = "r")

```

```{r}
#| include: false
# Some resources
# https://brunaw.com/slides/rladies-dublin/RF/intro-to-rf.html#27
# https://www.rebeccabarter.com/blog/2020-03-25_machine_learning#what-is-tidymodels
# https://juliasilge.com/blog/ikea-prices/
# https://juliasilge.com/blog/sf-trees-random-tuning/
# https://www.tidymodels.org/start/tuning/
# https://workflows.tidymodels.org/reference/predict-workflow.html
# https://christophm.github.io/interpretable-ml-book/feature-importance.html 
```

## Biomass trends with trends-covariates
First we split the data into training and testing (for evaluating the final model's performance after tuning hyper-parameters using cross validation)

```{r}
d |> 
  pivot_longer(c(temp_slope, mean_temp, oxy_slope, mean_oxy, sal_slope, mean_sal, mean_biom)) |> 
  ggplot(aes(value)) + 
  facet_wrap(~name, scales = "free") +
  geom_histogram()

trend <- d
```

```{r}
trend_split <- initial_split(trend, prop = 3/4)
trend_split

# Extract training and testing sets
trend_train <- training(trend_split)
trend_test <- testing(trend_split)

# We need to tune the model by comparing the predictive performance of hyper parameters using cross validation, so set up a cross-validation object.

# Create CV object from training data
trend_cv <- vfold_cv(trend_train)

# Now define the "recipe", basically the formula because we have already pre-processed the data
# Define the recipe, note the variables are already scaled
trend_recipe <- recipe(biom_trend ~ temp_slope + mean_temp + oxy_slope + mean_oxy + sal_slope + mean_sal + mean_biom,
                       data = trend)

trend_recipe

# Next step is to pre-specify the model using `parsnip` It's basically a "universal" way to specify a model by defining the model type, arguments (such as which hyper parameters to tune), engine. We tune mtry (number of predictors that are sampled at splits in a tree-based model node) and trees (how many models to average predictions over)

rf_trend_model <- 
  # specify that the model is a random forest
  rand_forest() |>
  # specify what hyperparameters to tune
  set_args(mtry = tune(),
           trees = tune()) |>
  # select the engine/package that underlies the model
  set_engine("ranger", importance = "permutation", keep.inbag = TRUE) |> # keep.inbag is to get CI's later!
  set_mode("regression")

rf_trend_model

# Put the model and recipes together into a workflow. You initiate a workflow using `workflow()`, and then you can add a recipe and add a model to it.

# Set the workflow
rf_trend_workflow <- workflow() |>
  add_recipe(trend_recipe) |>
  add_model(rf_trend_model)

rf_trend_workflow
```

Tune hyper parameters by checking model performance using cross validation

```{r}
# Default mtry is the (rounded down) square root of the number variables and tree is 500. Specify which values want to try in a grid
# rf_trend_grid <- expand.grid(
#   mtry = seq(2, 5, by = 1),
#   trees = seq(200, 1000, by = 200)
#   )
 
rf_trend_grid <- expand.grid(
  mtry = 5,
  trees = 600
  )

# Extract results of tuning (metric is rmse)
rf_trend_tune_results <- rf_trend_workflow |>
  tune_grid(resamples = trend_cv, # CV object
            grid = rf_trend_grid) # grid of values to try

# Print results
rf_trend_tune_results |>
  collect_metrics()

# autoplot(rf_trend_tune_results) +
#   scale_color_brewer(palette = "Set1") + 
#   theme(legend.position = "bottom")

rf_trend_tune_results |>
  collect_metrics() |>
  filter(.metric == "rmse") |> 
  ggplot(aes(x = trees, y = mean, color = as.factor(mtry))) +
  geom_point(alpha = 0.5) +
  geom_line() +
  labs(x = "Trees", y = "RMSE") +
  scale_color_brewer(palette = "Set1", 
                     name = "# Randomly\nselected\npredictors")

ggsave(paste0(home, "/figures/supp/tuning.pdf"), width = 12, height = 9, units = "cm")

# Now update the workflow to get the best mtry and n trees from the tuning
param_trend_final <- rf_trend_tune_results |>
  select_best(metric = "rmse")

param_trend_final

# Add this parameter to the workflow using the finalize_workflow() function.
rf_trend_workflow <- rf_trend_workflow |>
  finalize_workflow(param_trend_final) # can do the select best in here as well

rf_trend_workflow

# Now we’ve defined our recipe, our model, and tuned the model’s parameters, we can fit the final model. Applying the `last_fit()` function to our workflow and our train/test split object will train the model specified by the workflow using the training data, and produce evaluations based on the test set.

rf_trend_fit <- rf_trend_workflow |>
  last_fit(trend_split)

# Evaluate the model performance on the testing data

rf_trend_fit |> collect_metrics()

# Now that we have a final model we want to train it on the full dataset, and then use it to predict the response for new data. For that we need to `fit()` again on the full dataset (the complete training + testing dataset).

final_trend_model <- fit(rf_trend_workflow, trend)

final_trend_model

# Save!
saveRDS(final_trend_model, paste0(home, "/output/rf_final_trend_model.rds"))
```

Fit plots

```{r}
# Plot results

collect_predictions(rf_trend_fit) |>
  ggplot(aes(biom_trend, .pred)) +
  geom_point(alpha = 0.5, color = "steelblue") +
  labs(x = "Observed", y = "Predicted") +
  geom_abline(linetype = 2, color = "tomato3") +
  NULL

ggsave(paste0(home, "/figures/supp/obs_pred.pdf"), width = 11, height = 11, units = "cm")
```

Load model

```{r}
final_trend_model <- readRDS(paste0(home, "/output/rf_final_trend_model.rds"))
```

Look at variable importance

```{r trend vip}
#| message: false
# p <- 
#   # TODO: or trend_train?
#   final_trend_model_old |> 
#   extract_fit_parsnip() |>
#   vip(aesthetics = list(alpha = 0.8, fill = "steelblue"))

# For a tailor made plot we can also extract the VIPs
trend_ranger_obj <- extract_fit_parsnip(final_trend_model)$fit
#trend_ranger_obj$variable.importance

create_vip_tibble <- function(trend_ranger_obj) {
  tibble(Importance = trend_ranger_obj$variable.importance,
         var = as.factor(names(trend_ranger_obj$variable.importance))) |> 
    mutate(var2 = recode(var, 
                         "mean_biom" = "Mean biomass density",
                         "oxy_slope" = "Oxygen velocity",
                         "sal_slope" = "Salinity slope",
                         "temp_slope" = "Temperature slope",
                         "mean_temp" = "Mean temperature",
                         "mean_sal" = "Mean salinity",
                         "mean_oxy" = "Mean oxygen",
                         "group" = "Group"),
           response = "Trend")
}

vip_tib_trend <- create_vip_tibble(trend_ranger_obj)

p1 <- ggplot(vip_tib_trend,
               aes(Importance, reorder(var2, Importance))) + 
  #geom_point() + 
  #geom_segment(aes(x = 0, xend = Importance,
  #                 yend = as.numeric(reorder(var2, Importance)))) +
  geom_bar(stat = "identity", fill = "grey40") +
  coord_cartesian(expand = 0) +
  theme(axis.title.y = element_blank()) +
  annotate("text", label = "(a)", x = -Inf, y = Inf, vjust = 1.5, hjust = -0.5, color = "white", size = 3)

# Partial dependence plot
# https://stackoverflow.com/questions/67634344/r-partial-dependence-plots-from-workflow
explainer_rf <- explain_tidymodels(
    final_trend_model, 
    data = dplyr::select(trend, -biom_trend), 
    y = trend$biom_trend,
    label = "random forest"
)

acc_rf <- model_profile(
  explainer_rf,
  variables = "mean_temp",
  type = "accumulated"
)

plot(acc_rf)

acc_rf <- model_profile(
  explainer_rf,
  variables = "mean_biom",
  type = "accumulated"
)

plot(acc_rf)

pdp_rf <- model_profile(explainer_rf, N = NULL, 
                        variables = c(
                          "mean_temp",
                          # "mean_oxy",
                          # "mean_sal",
                          # "temp_slope",
                          # "oxy_slope",
                          # "sal_slope",
                          "mean_biom"
                          ),
                        groups = NULL)

plot(pdp_rf)

pd <- as_tibble(pdp_rf$agr_profiles) |> 
  janitor::clean_names() |> 
  mutate(var2 = recode(vname,
                       "mean_biom" = "Mean biomass density",
                       "oxy_slope" = "Oxygen slope",
                       "temp_slope" = "Temperature slope",
                       "sal_slope" = "Salinity slope",
                       "mean_temp" = "Mean temperature (°C)",
                       "mean_oxy" = "Mean oxygen (ml/L)",
                       "mean_sal" = "Mean salinity (psu)",
                       "Group" = "Group")) |> 
  mutate(group2 = str_replace(groups, "_", " "),
         group2 = str_to_title(group2),
         group2 = factor(group2, 
                         levels = c("Cod Adult", "Cod Juvenile", "Flounder Adult",
                                    "Flounder Juvenile",
                                    "Plaice Adult", "Plaice Juvenile", "Dab All"))) |> 
  mutate(yhat_ct = (yhat - mean(yhat)) / sd(yhat), .by = group2)

facet_labels <- tibble(
  var2 = c("Mean temperature (°C)", "Mean biomass density"),
  label = c("(c)", "(b)"),
  x = -Inf,
  y = Inf
)

p2 <- pd |> 
  filter(var2 %in% c("Mean temperature (°C)", "Mean biomass density")) |> 
  #filter((between(x, c(-2, 2)) & var2 == "Mean biomass density")) |> 
  ggplot(aes(x, yhat_ct, color = group2)) +
  facet_wrap(~var2, scales = "free") +
  geom_line(alpha = 0.8, linewidth = 0.9) +
  scale_color_brewer(palette = "Paired") +
  geom_text(data = facet_labels,
            aes(x = x, y = y, label = label),
            inherit.aes = FALSE,
            hjust = -0.5, 
            vjust = 1.5,
            color = "gray40",
            size = 3) +
  labs(x = "Predictor value",
       y = "z-scored predicted biomass trend") +
  geom_smooth(aes(x, yhat_ct), inherit.aes = FALSE, color = "gray40", 
              alpha = 0.4, linewidth = 0.75,
              method = "gam", formula = y~s(x, k=3)) +
  guides(color = guide_legend(hjust = -10)) +
  theme(legend.position = "bottom",
        legend.title = element_blank())

(free(p1) / p2) + plot_layout(heights = c(1, 1.3))

ggsave(paste0(home, "/figures/vip_effects.pdf"), width = 18, height = 16, units = "cm")
```

```{r}

rf_trend_fit |> collect_metrics()
```


```{r}
part_dep_all <- pd |>
  ggplot(aes(x, yhat_ct, color = group2)) +
  geom_line(alpha = 0.8, linewidth = 0.9) +
  facet_wrap(~var2, ncol = 3, scales = "free") +
  scale_color_brewer(palette = "Paired") +
  labs(x = "Predictor value",
       y = "Predicted biomass trend") +
  geom_smooth(aes(x, yhat_ct), inherit.aes = FALSE, color = "gray40", 
              alpha = 0.4, linewidth = 0.75,
              method = "gam", formula = y~s(x, k=3)) +
  guides(color = guide_legend(position = "inside")) +
  theme(legend.position.inside = c(0.85, 0.25),
        legend.title = element_blank())

part_dep_all
 
ggsave(paste0(home, "/figures/supp/partial_dependence.pdf"), width = 18, height = 11, units = "cm")
```

```{r}
knitr::knit_exit()
```

```{r}
## Plot conditional effects and partial dependence plots

# Conditional effect of temperature (using the range in the old data)
# nd_temp_old <- trend_old |>
#   group_by(group) |> 
#   data_grid(mean_temp = seq_range(mean_temp, n = 100, trim = 0.1)) |> 
#   ungroup() |> 
#   mutate(temp_slope = mean(d_old$temp_slope),
#          oxy_slope = mean(d_old$oxy_slope),
#          mean_biom = mean(d_old$mean_biom),
#          mean_oxy = mean(d_old$mean_oxy),
#          group2 = str_replace(group, "_", " "),
#          group2 = str_to_sentence(group2)) |> 
#   separate_wider_delim(group2, delim = " ", names = c("Species", "Life stage"), 
#                        cols_remove = FALSE) |> 
#   mutate(`Life stage` = str_to_sentence(`Life stage`))
# 
# # Predict
# nd_temp_old$pred <- predict(final_trend_model_old, new_data = nd_temp_old)$.pred
# 
# # Predict and get jacknife CI's
# # So, we don't have a ranger object but a workflow object... 
# # https://workflows.tidymodels.org/reference/predict-workflow.html
# # I assume because this is a ranger model, the se.method = "infjack"
# # https://github.com/imbs-hl/ranger/issues/136
# pred_ci_old <- predict(final_trend_model_old, new_data = nd_temp_old, type = "conf_int", level = 0.95)
# 
# nd_temp_old <- nd_temp_old |> bind_cols(pred_ci_old)
#   
# 
# # Now the shorter time frame
# nd_temp_new <- trend_new |>
#   group_by(group) |> 
#   data_grid(mean_temp = seq_range(mean_temp, n = 100, trim = 0.1)) |> 
#   ungroup() |> 
#   mutate(temp_slope = mean(d_new$temp_slope),
#          oxy_slope = mean(d_new$oxy_slope),
#          mean_biom = mean(d_new$mean_biom),
#          mean_oxy = mean(d_new$mean_oxy),
#          group2 = str_replace(group, "_", " "),
#          group2 = str_to_sentence(group2)) |> 
#   separate_wider_delim(group2, delim = " ", names = c("Species", "Life stage"), 
#                        cols_remove = FALSE) |> 
#   mutate(`Life stage` = str_to_sentence(`Life stage`))
#   
# nd_temp_new$pred <- predict(final_trend_model_new, new_data = nd_temp_new)$.pred
# 
# pred_ci_new <- predict(final_trend_model_new, new_data = nd_temp_new, type = "conf_int", level = 0.95)
# 
# nd_temp_new <- nd_temp_new |> bind_cols(pred_ci_new)
# 
# nd_temp <- bind_rows(nd_temp_old, nd_temp_new) |> 
#   mutate(pred_ct = pred - mean(pred),
#          .pred_lower_ct = .pred_lower - mean(.pred_lower),
#          .pred_upper_ct = .pred_upper - mean(.pred_upper),
#          .by = group)
# 
# hist(nd_temp$pred)
# 
# pal <- brewer.pal(n = 12, name = "Paired")[c(1:6, 8, 10, 12)]
# 
# # Plot conditional effect of temperature
# cond_temp <- ggplot(nd_temp, aes(mean_temp, pred_ct, color = group2)) + 
#   geom_ribbon(aes(ymin = .pred_lower_ct, ymax = .pred_upper_ct, fill = group2),
#               alpha = 0.15, color = NA) +
#   geom_line(alpha = 0.8, linewidth = 0.9) + 
#   facet_wrap(~"Conditional effect") +
#   scale_color_manual(values = pal) +
#   scale_fill_manual(values = pal) +
#   labs(x = "Mean temperature (°C)", 
#        y = "Scaled local biomass trend") + 
#   geom_smooth(aes(mean_temp, pred_ct), inherit.aes = FALSE,
#               color = "black", alpha = 0.4, size = 0.75) +
#   annotate("text", label = "(c)", x = -Inf, y = Inf, vjust = 1.5, hjust = -0.2, color = "gray40", size = 4) +
#   guides(fill = "none",
#           color = "none")
# 
# cond_temp
```

