---
title: "Bias-correcting environmental data"
author: "Max Lindmark"
date: "`r format(Sys.time(), '%d %B, %Y')`"
format: 
  html:
    embed-resources: true
    fig-width: 8
knitr: 
  opts_chunk:
    fig.align: center
editor: source
execute: 
  echo: true
  eval: true
  cache: false
---

Read in hindcast and new environmental data and perform bias correction on the new data.

```{r setup}
#| warning: false
#| message: false

# Load libraries, install if needed
library(tidyverse)
library(tidync)
library(tidyterra)
library(ncdf4)
library(viridis)
library(sdmTMB)
library(terra)
library(patchwork)
library(ggsidekick); theme_set(theme_sleek())
library(tidylog)

# Set path
home <- here::here()
```

## Read data

```{r}
#| warning: false
#| message: false

# Oxygen
# https://data.marine.copernicus.eu/product/BALTICSEA_MULTIYEAR_BGC_003_012/download?dataset=cmems_mod_bal_bgc_my_P1M-m_202303
# Print details
print(nc_open(paste0(home, "/data/covariates/cmems_mod_bal_bgc_my_P1M-m_1742291088556.nc")))

oxy_tibble <- tidync(paste0(home, "/data/covariates/cmems_mod_bal_bgc_my_P1M-m_1742291088556.nc")) |>
  hyper_tibble() |>
  mutate(date = as_datetime(time, origin = "1970-01-01")) |>
  mutate(
    month = month(date),
    day = day(date),
    year = year(date),
    oxy = o2b * 0.0223909
  )

# Salinity and temperature
# https://data.marine.copernicus.eu/product/BALTICSEA_MULTIYEAR_PHY_003_011/download?dataset=cmems_mod_bal_phy_my_P1M-m_202303
# Print details
print(nc_open(paste0(home, "/data/covariates/cmems_mod_bal_phy_my_P1M-m_1742290961273.nc")))

st_tibble <- tidync(paste0(home, "/data/covariates/cmems_mod_bal_phy_my_P1M-m_1742290961273.nc")) |>
  hyper_tibble() |>
  mutate(date = as_datetime(time, origin = "1970-01-01")) |>
  mutate(
    month = month(date),
    day = day(date),
    year = year(date)
  )

# Combine old environmental data
coper <- st_tibble |> 
  left_join(oxy_tibble |> dplyr::select(longitude, latitude, oxy, year, month),
            by = c("longitude", "latitude", "year", "month"))

coper <- coper |> 
  rename(temp = bottomT,
         lon = longitude, 
         lat = latitude,
         sal = sob) |> 
  dplyr::select(-time, -date, -day) |> 
  mutate(source = "coper",
         year_month = paste(year, month, sep = "_"))

# Hindcast
hind <- readRDS(file = paste0(home, "/data/covariates/hindcast_1961_2017.rds")) |> 
  dplyr::select(-hypoxia1, -hypoxia2, -anoxia, -time, -depth) |> 
  mutate(source = "hindcast",
         year_month = paste(year, month, sep = "_")) |> 
  filter(year >= 1993) |> # This is the year where copernicus data start
  filter(lat <= max(coper$lat))

extra_time <- hind |>
  filter(year == max(year)) |>
  mutate(across(c(temp, sal, oxy, year_month), ~NA)) |>
  replicate_df("year", setdiff(coper$year, hind$year)) |> 
  mutate(year_month = paste(year, month, sep = "_")) |> 
  filter(year_month %in% unique(coper$year_month))

hind <- hind |> bind_rows(extra_time)
```

## Extract values of copernicus to hindcast locations

The two sources are not on the same grid. For fitting, just bind-row then and add a source column. For a wide data, extract with terra.

```{r}
#| warning: false
#| message: false

# Loop through all year combos, extract the copernicus temperatures at the data locations of the hindcast
env_list <- list()

for (i in unique(hind$year_month)) {

  d_sub <- filter(hind, year_month == i)
  env_sub <- filter(coper, year_month == i)

  # Convert to spat_raster
  env_raster <- as_spatraster(env_sub,
    xycols = 3:4,
    crs = "WGS84", digits = 3
  )

  # Extract from raster
  d_sub$temp_n <- terra::extract(
    env_raster$temp,
    d_sub |> dplyr::select(lon, lat)
  )$temp

  d_sub$sal_n <- terra::extract(
    env_raster$sal,
    d_sub |> dplyr::select(lon, lat)
  )$sal

  d_sub$oxy_n <- terra::extract(
    env_raster$oxy,
    d_sub |> dplyr::select(lon, lat)
  )$oxy

  # Save
  env_list[[i]] <- d_sub
}

wide <- bind_rows(env_list)
```

```{r}
# TODO: truncate 0; negative means oxygen is depleted
wide |> 
  filter(year > 2000 & year < 2017) |> 
  pivot_longer(c(oxy, oxy_n)) |> 
  ggplot(aes(value, fill = name)) + 
  geom_density(alpha = 0.4)

wide |> 
  filter(year > 2000 & year < 2017) |> 
  pivot_longer(c(oxy, oxy_n)) |> 
  summarise(minvalue = min(value, na.rm = TRUE), .by = name)
```

Read in old hindcast data (prior to copernicus)

```{r}
#| warning: false
#| message: false

hind_old <- readRDS(file = paste0(home, "/data/covariates/hindcast_1961_2017.rds")) |> 
  dplyr::select(-hypoxia1, -hypoxia2, -anoxia, -time, -depth) |> 
  mutate(year_month = paste(year, month, sep = "_")) |> 
  filter(year < 1993 & year >= 1978) |> 
  filter(lat <= max(coper$lat)) |> 
  mutate(sal_n = NA,
         oxy_n = NA,
         temp_n = NA,
         source = "hindcast") |> 
  filter(lon < 22) |> 
  filter(!(lat > 55.5 & lon < 13))

wide <- bind_rows(wide, hind_old) |> 
  filter(lon < 22) |> 
  filter(!(lat > 55.5 & lon < 13))

wide <- wide |>
  add_utm_columns(ll_names = c("lon", "lat"))
```

## Bias correcting with difference- model

### Oxygen

```{r}
#| warning: false
#| message: false

oxy_dat <- wide |> 
  filter(year %in% c(2001:2017)) |> 
  mutate(diff = oxy - oxy_n) |> 
  drop_na(diff)

# Make mesh & fit models
mesh <- make_mesh(oxy_dat, cutoff = 10, xy_cols = c("X", "Y"))
#plot(mesh)

proj_years <- setdiff(coper$year, drop_na(hind)$year)
hind_year <- min(coper$year):(min(oxy_dat$year)-1)

# This assumes seasonality is the same.
# It does capture spatial differences between the models, and the fact that the mean difference seems to vary from year to year. We do this with a random walk, so the difference in means in years 2017--2024 is basically a projection... We can use 1993--2006 for comparing our predictions (then we need extra time I think)
moxy <- sdmTMB(diff ~ 1,
               spatial = "on",
               time_varying = ~1,
               time_varying_type = "rw0",
               extra_time = c(hind_year, proj_years),
               mesh = mesh,
               time = "year",
               data = oxy_dat
               )

moxy
sanity(moxy)

# Time-varying effect
est <- as.list(moxy$sd_report, "Est")
est_se <- as.list(moxy$sd_report, "Std. Error")

tw <- tibble(est = est$b_rw_t,
             se = est_se$b_rw_t,
             year = c(hind_year, unique(oxy_dat$year), proj_years))

ptw <- ggplot(tw, aes(year, est)) + 
  geom_ribbon(aes(ymin = est - 1.96*se, ymax = est + 1.96*se),
              alpha = 0.3) +
  labs(x = "Year", y = "Time-varying intercept") +
  coord_cartesian(expand = 0) +
  xlim(1978, 2024) +
  annotate("text", label = "(a)", x = -Inf, y = Inf, vjust = 1.5, hjust = -0.2,
           color = "gray40", size = 4) +
  geom_line() 

# Predict differences (including the extra_time)
poxy <- predict(moxy, 
                wide |> filter(year %in% c(hind_year, unique(oxy_dat$year), proj_years)))

# Add in the pre-copernicus years
poxy <- bind_rows(poxy, hind_old)

# Manually predict the hindcast values, given copernicus and the estimated difference
poxy$oxy_pred <- poxy$oxy_n + poxy$est

# Time series plots
pal <- RColorBrewer::brewer.pal(n=4, "Paired")
 
pmean <- poxy |> 
  pivot_longer(c(oxy, oxy_n, oxy_pred)) |> 
  mutate(name = fct_recode(name,
    "Hindcast" = "oxy",
    "Copernicus" = "oxy_n",
    "Hindcast-predicted" = "oxy_pred"
  )) |>
  summarise(mean = mean(value, na.rm = TRUE), .by = c(name, year)) |> 
  ggplot(aes(year, mean, color = name, linetype = name)) + 
  geom_rect(aes(xmin = min(oxy_dat$year), xmax = max(proj_years),
                ymin = -Inf, ymax = Inf), color = NA,
              fill = "grey95", alpha = 0.1) +
  geom_rect(aes(xmin = min(proj_years), xmax = max(proj_years),
                ymin = -Inf, ymax = Inf), color = NA,
              fill = "grey85", alpha = 0.1) +
  geom_line(linewidth = 1.1) +
  scale_color_manual(values = pal[c(1,4,2)]) +
  labs(x = "Year", y = expression(paste("Mean O"[2], " (ml/L)"))) +
  theme(legend.position.inside = c(0.15, 0.15),
        legend.title = element_blank()) + 
  guides(color = guide_legend(position = "inside")) +
  coord_cartesian(expand = 0) +
  annotate("text", label = "(b)", x = -Inf, y = Inf, vjust = 1.5, hjust = -0.2,
           color = "gray40", size = 4)

(ptw / pmean) + plot_layout(axes = "collect")

ggsave(paste0(home, "/figures/supp/bias_corr_oxy_means.pdf"), width = 17, height = 14, units = "cm")
```

### Temperature

```{r}
#| warning: false
#| message: false

temp_dat <- wide |> 
  filter(year %in% c(2001:2017)) |> 
  mutate(diff = temp - temp_n) |> 
  drop_na(diff)

# Make mesh & fit models
mesh <- make_mesh(temp_dat, cutoff = 10, xy_cols = c("X", "Y"))
#plot(mesh)

# Fit
mtemp <- sdmTMB(diff ~ 1,
                spatial = "on",
                time_varying = ~1,
                time_varying_type = "rw0",
                extra_time = c(hind_year, proj_years),
                mesh = mesh,
                time = "year",
                data = temp_dat
                )

mtemp
sanity(mtemp)

# Time-varying effect
est <- as.list(mtemp$sd_report, "Est")
est_se <- as.list(mtemp$sd_report, "Std. Error")

tw <- tibble(est = est$b_rw_t,
             se = est_se$b_rw_t,
             year = c(hind_year, unique(temp_dat$year), proj_years))

ptw <- ggplot(tw, aes(year, est)) + 
  geom_ribbon(aes(ymin = est - 1.96*se, ymax = est + 1.96*se),
              alpha = 0.3) +
  labs(x = "Year", y = "Time-varying intercept") +
  coord_cartesian(expand = 0) +
  xlim(1978, 2024) +
  annotate("text", label = "(a)", x = -Inf, y = Inf, vjust = 1.5, hjust = -0.2,
           color = "gray40", size = 4) +
  geom_line() 

# Predict differences (including the extra_time)
ptemp <- predict(mtemp, wide |> filter(year %in% c(hind_year, unique(temp_dat$year), proj_years)))

# Add in the pre-copernicus years
ptemp <- bind_rows(ptemp, hind_old)

# Manually predict the hindcast values, given copernicus and the estimated difference
ptemp$temp_pred <- ptemp$temp_n + ptemp$est

# Time series plots
pal <- RColorBrewer::brewer.pal(n=4, "Paired")
 
pmean <- ptemp |> 
  pivot_longer(c(temp, temp_n, temp_pred)) |> 
  mutate(name = fct_recode(name,
    "Hindcast" = "temp",
    "Copernicus" = "temp_n",
    "Hindcast-predicted" = "temp_pred"
  )) |>
  summarise(mean = mean(value, na.rm = TRUE), .by = c(name, year)) |> 
  ggplot(aes(year, mean, color = name, linetype = name)) + 
  geom_rect(aes(xmin = min(temp_dat$year), xmax = max(proj_years),
                ymin = -Inf, ymax = Inf), color = NA,
              fill = "grey95", alpha = 0.1) +
  geom_rect(aes(xmin = min(proj_years), xmax = max(proj_years),
                ymin = -Inf, ymax = Inf), color = NA,
              fill = "grey85", alpha = 0.1) +
  geom_line(linewidth = 1.1) +
  scale_color_manual(values = pal[c(1,4,2)]) +
  labs(x = "Year", y = "Mean temperature (°C)") +
  theme(legend.position.inside = c(0.15, 0.85),
        legend.title = element_blank()) + 
  guides(color = guide_legend(position = "inside")) +
  coord_cartesian(expand = 0) +
  annotate("text", label = "(b)", x = -Inf, y = Inf, vjust = 1.5, hjust = -0.2,
           color = "gray40", size = 4)

(ptw / pmean) + plot_layout(axes = "collect")

ggsave(paste0(home, "/figures/supp/bias_corr_temp_means.pdf"), width = 17, height = 14, units = "cm")
```

### Salinity

```{r}
#| warning: false
#| message: false

sal_dat <- wide |> 
  filter(year %in% c(2001:2017)) |> 
  mutate(diff = sal - sal_n) |> 
  drop_na(diff)

# Make mesh & fit models
mesh = make_mesh(sal_dat, cutoff = 10, xy_cols = c("X", "Y"))
#plot(mesh)

# Fit
msal <- sdmTMB(diff ~ 1,
               spatial = "on",
               time_varying = ~1,
               time_varying_type = "rw0",
               extra_time = c(hind_year, proj_years),
               mesh = mesh,
               time = "year",
               data = sal_dat
               )

msal
sanity(msal)

# Time-varying effect
est <- as.list(msal$sd_report, "Est")
est_se <- as.list(msal$sd_report, "Std. Error")

tw <- tibble(est = est$b_rw_t,
             se = est_se$b_rw_t,
             year = c(hind_year, unique(sal_dat$year), proj_years))

ptw <- ggplot(tw, aes(year, est)) + 
  geom_ribbon(aes(ymin = est - 1.96*se, ymax = est + 1.96*se),
              alpha = 0.3) +
  labs(x = "Year", y = "Time-varying intercept") +
  coord_cartesian(expand = 0) +
  xlim(1978, 2024) +
  annotate("text", label = "(a)", x = -Inf, y = Inf, vjust = 1.5, hjust = -0.2,
           color = "gray40", size = 4) +
  geom_line() 

# Predict differences (including the extra_time)
psal <- predict(msal, wide |> filter(year %in% c(hind_year, unique(sal_dat$year), proj_years)))

# Add in the pre-copernicus years
psal <- bind_rows(psal, hind_old)

# Manually predict the hindcast values, given copernicus and the estimated difference
psal$sal_pred <- psal$sal_n + psal$est

# Time series plots
pal <- RColorBrewer::brewer.pal(n=4, "Paired")
 
pmean <- psal |> 
  pivot_longer(c(sal, sal_n, sal_pred)) |> 
  mutate(name = fct_recode(name,
    "Hindcast" = "sal",
    "Copernicus" = "sal_n",
    "Hindcast-predicted" = "sal_pred"
  )) |>
  summarise(mean = mean(value, na.rm = TRUE), .by = c(name, year)) |> 
  ggplot(aes(year, mean, color = name, linetype = name)) + 
  geom_rect(aes(xmin = min(temp_dat$year), xmax = max(proj_years),
                ymin = -Inf, ymax = Inf), color = NA,
              fill = "grey95", alpha = 0.1) +
  geom_rect(aes(xmin = min(proj_years), xmax = max(proj_years),
                ymin = -Inf, ymax = Inf), color = NA,
              fill = "grey85", alpha = 0.1) +
  geom_line(linewidth = 1.1) +
  scale_color_manual(values = pal[c(1,4,2)]) +
  labs(x = "Year", y = "Mean salinity (ppt)") +
  theme(legend.position.inside = c(0.15, 0.85),
        legend.title = element_blank()) + 
  guides(color = guide_legend(position = "inside")) +
  coord_cartesian(expand = 0) +
  annotate("text", label = "(b)", x = -Inf, y = Inf, vjust = 1.5, hjust = -0.2,
           color = "gray40", size = 4)

(ptw / pmean) + plot_layout(axes = "collect")

ggsave(paste0(home, "/figures/supp/bias_corr_sal_means.pdf"), width = 17, height = 14, units = "cm")
```

## Make spatial plots

```{r}
#| warning: false
#| message: false

# Plot environmental data for example year
poxys <- filter(poxy, year == 2012 & month == 10) |> 
  rename(Hindcast = oxy, 
         Copernicus = oxy_n,
         `Predicted Hindcast` = oxy_pred) |> 
  pivot_longer(c("Hindcast", "Copernicus", "Predicted Hindcast")) |> 
  ggplot() +
  geom_raster(aes(lon, lat, fill = value)) + 
  facet_wrap(~name) +
  scale_fill_viridis(option = "mako", name = expression(paste("O"[2], " (ml/L)"))) + 
  guides(fill = guide_colorbar(position = "inside")) +
  labs(x = "Longitude", y = "Latitude") +
  theme(legend.position.inside = c(0.052, 0.73),
        legend.key.size = unit(0.25, "cm"),
        axis.title.x = element_blank(),
        axis.title.y = element_blank(),
        axis.text.x = element_blank())

ptemps <- filter(ptemp, year == 2012 & month == 10) |> 
  rename(Hindcast = temp, 
         Copernicus = temp_n,
         `Predicted Hindcast` = temp_pred) |> 
  pivot_longer(c("Hindcast", "Copernicus", "Predicted Hindcast")) |> 
  ggplot() +
  geom_raster(aes(lon, lat, fill = value)) + 
  facet_wrap(~name) +
  scale_fill_viridis(option = "magma", name = "Temperature\n(°C)") + 
  guides(fill = guide_colorbar(position = "inside")) +
  labs(x = "Longitude", y = "Latitude") +
  theme(legend.position.inside = c(0.07, 0.73),
        strip.text.x = element_blank(),
        legend.key.size = unit(0.25, "cm"),
        axis.title.x = element_blank(),
        axis.text.x = element_blank())

psals <- filter(psal, year == 2012 & month == 10) |> 
  rename(Hindcast = sal, 
         Copernicus = sal_n,
         `Predicted Hindcast` = sal_pred) |> 
  pivot_longer(c("Hindcast", "Copernicus", "Predicted Hindcast")) |> 
  ggplot() +
  geom_raster(aes(lon, lat, fill = value)) + 
  facet_wrap(~name) +
  scale_fill_viridis(option = "D", name = "Salinity (ppt)") + 
  guides(fill = guide_colorbar(position = "inside")) +
  labs(x = "Longitude", y = "Latitude") +
  theme(legend.position.inside = c(0.068, 0.73),
        strip.text.x = element_blank(),
        legend.key.size = unit(0.25, "cm"),
        axis.title.y = element_blank())

(poxys / ptemps / psals)

ggsave(paste0(home, "/figures/supp/bias_corr_maps.pdf"), width = 18, height = 18, units = "cm")
```

## Save data

```{r}
#| warning: false
#| message: false

save_oxy <- poxy |> 
  filter(source == "hindcast") |> 
  mutate(oxy_full = oxy,
         oxy_full = ifelse(is.na(oxy_full), oxy_pred, oxy_full)) |> 
  dplyr::select(lon, lat, year, month, oxy_full) |> 
  rename(oxy = oxy_full) |> 
  mutate(id = paste(lon, lat, year, month, sep = "_")) |> 
  distinct(id, .keep_all = TRUE) |> 
  arrange(year, month)

save_temp <- ptemp |> 
  filter(source == "hindcast") |> 
  mutate(temp_full = temp,
         temp_full = ifelse(is.na(temp_full), temp_pred, temp_full)) |> 
  dplyr::select(lon, lat, year, month, temp_full) |> 
  rename(temp = temp_full) |> 
  mutate(id = paste(lon, lat, year, month, sep = "_")) |> 
  distinct(id, .keep_all = TRUE) |> 
  arrange(year, month)

save_sal <- psal |> 
  filter(source == "hindcast") |> 
  mutate(sal_full = sal,
         sal_full = ifelse(is.na(sal_full), sal_pred, sal_full)) |> 
  dplyr::select(lon, lat, year, month, sal_full) |> 
  rename(sal = sal_full) |> 
  mutate(id = paste(lon, lat, year, month, sep = "_")) |> 
  distinct(id, .keep_all = TRUE) |> 
  arrange(year, month)

bias_corr_vars <- save_oxy |> 
  left_join(save_temp |> dplyr::select(id, temp), by = "id") |> 
  left_join(save_sal |> dplyr::select(id, sal), by = "id")
  
# Save only grids that I have temperature in for all years (since I don't have hindcast between 2018--2023, there are a few cells with NA predicted variables)

grid_keep <- bias_corr_vars |> 
  filter(year == 2022) |> 
  mutate(id = paste(lon, lat, sep = "_")) |> 
  drop_na(oxy) |> pull(id)

bias_corr_vars <- bias_corr_vars |> 
  mutate(id = paste(lon, lat, sep = "_")) |> 
  filter(id %in% grid_keep) |>
  dplyr::select(-id)

write_csv(bias_corr_vars, paste0(home, "/data/covariates/bias_corr_vars.csv"))
```
