---
title: "Fit models to biomass density and extract AIC"
author: "Max Lindmark"
date: today
date-format: iso
toc: true
format: 
  html:
    page-layout: full
    embed-resources: true
knitr: 
  opts_chunk:
    fig.align: center
    out-width: 100%
editor: source
---

## Intro

Fit models to biomass density of cod, flounder, plaice and dab (juveniles and adults) between 1993-2023 in the Baltic Sea using sdmMTB fit different oxygen and temperature-related covariates, compare AIC. Select the "best" covariate for further trend and velocity analysis.

## Load packages & source functions

```{r load libraries}
#| message: false
#| warning: false

library(tidyverse)
library(tidylog)
library(readxl)
library(readr)
library(RCurl)
library(devtools)
library(patchwork)
library(kableExtra)
library(viridis)
library(RColorBrewer)
library(sdmTMB)

# Source code for map plots
devtools::source_url("https://raw.githubusercontent.com/maxlindmark/spatial-metabolic-index/main/R/functions/map-plot.R")
options(ggplot2.continuous.colour = "viridis")

# Set path
home <- here::here()
```

## Read data

```{r read data}
# Read data
d <- read_csv(paste0(home, "/data/clean/catch_clean.csv")) |>
  add_utm_columns(ll_names = c("lon", "lat")) |> 
  mutate(life_stage = ifelse(str_ends(group, "juvenile"), "juvenile", "all"),
         life_stage = ifelse(str_ends(group, "adult"), "adult", life_stage))

# Read metabolic parameter estimates and left_join
mi_pars <- read_csv(paste0(home, "/data/clean/mi_params.csv")) 

mi_pars <- mi_pars |> 
  rename(common_name = species) |> 
  mutate(A_0 = ifelse(common_name == "flounder",
                      filter(mi_pars, common_name == "plaice")$A_0,
                      A_0))

d <- left_join(d, mi_pars, by = "common_name")

# Read size csv to calculate the metabolic index
sizes <- read_csv(paste0(home, "/data/clean/sizes.csv")) |>
  mutate(group = paste(common_name, life_stage, sep = "_")) |> 
  dplyr::select(group, B, species)

d <- tidylog::left_join(d, sizes, by = "group")
```

## Calculate the metabolic index

```{r calculate metabolic index}
# Oxygen is ml/L, We want micro mol/L. 1 ml/l = 10^3/22.391 = 44.661 micro mol/l
d$oxy_si <- (d$oxy * (10^3)) / 22.391

# Calculate metabolic indices for a given mass and the temperature and oxygen in data
# Line 123 in https://github.com/fate-spatialindicators/SDM_O2/blob/master/code/mi_functions.R
kb <- 0.000086173324 # Boltzmann's constant
t_ref <- 15 # arbitrary reference temperature

# Calculate the metabolic index
d <- d |>
  mutate(inv_temp = (1/(temp + 273.15) - 1/(t_ref + 273.15)),
         phi = A_0*(B^n)*oxy_si * exp((E/kb)*inv_temp))
```

## Scale variables

```{r}
pred_grid <- bind_rows(readr::read_csv(paste0(home, "/data/clean/pred_grid_(1_2).csv")),
                       readr::read_csv(paste0(home, "/data/clean/pred_grid_(2_2).csv")))

d <- d |> 
  mutate(temp_sc = (temp - mean(pred_grid$temp)) / sd(pred_grid$temp),
         temp_sq = temp_sc^2,
         oxy_sc = (oxy - mean(pred_grid$oxy)) / sd(pred_grid$oxy),
         oxy_sq = oxy_sc^2,
         sal_sc = (sal - mean(pred_grid$sal)) / sd(pred_grid$sal),
         depth_sc = (depth - mean(pred_grid$depth)) / sd(pred_grid$depth),
         depth_sq = depth_sc*depth_sc,
         month_f = as.factor(month),
         year_f = as.factor(year)) |> 
  mutate(phi_sc = as.vector(scale(phi)), .by = group) # don't scale this across a common value because it varies by species

# Based on poor model fit due to infrequent observations, I drop these
d <- d |> filter(!group %in% c("whiting_all", "turbot_all"))
```

```{r}
order <- c("Cod Adult", "Cod Juvenile", "Flounder Adult", "Flounder Juvenile",
           "Plaice Adult", "Plaice Juvenile", "Dab All")

d |> 
  mutate(life_stage = str_to_sentence(life_stage),
         group2 = paste(str_to_sentence(common_name), str_to_sentence(life_stage)),
         group2 = factor(group2, levels = order)) |> 
  ggplot(aes(phi, fill = group2, color = group2)) + 
  geom_density(alpha = 0.2) + 
  geom_vline(xintercept = 1, linetype = 2) + 
  coord_cartesian(expand = 0) + 
  facet_wrap(~ life_stage, ncol = 1) +
  theme(legend.position = "bottom",
        legend.key.size = unit(0.4, "cm"),
        legend.title = element_blank()) +
  labs(x = "Metabolic index (\u03C6)", y = "Density", fill = "Species", color = "Species") + 
  scale_color_brewer(palette = "Paired") +
  scale_fill_brewer(palette = "Paired") +
  NULL

ggsave(paste0(home, "/figures/supp/mi_density.pdf"), width = 13, height = 16, units = "cm", device = cairo_pdf)
```

## Fit models and save marginal AIC

```{r fit models}
data_list_aic <- list()

for(i in unique(d$group)) {  
    
    dd <- d |> filter(group == i)
    
    # For some combos, there are years without a single occurrence. This makes it hard to fit year_f models...
    dd <- dd |> 
      mutate(sum = sum(density), .by = year) |> 
      filter(sum > 0)
    
    plot_map_fc + 
      geom_point(data = dd |> mutate(pres = ifelse(density == 0, "0", "1")),
                 aes(X*1000, Y*1000, color = pres)) +
      facet_wrap(~year)
    
    mesh <- make_mesh(dd, xy_cols = c("X", "Y"), cutoff = 12)
    
    print(
    ggplot() +
      inlabru::gg(mesh$mesh) +
      coord_fixed() +
      geom_point(aes(X, Y), data = dd, alpha = 0.2, size = 0.5) +
      labs(x = "Easting (km)", y = "Northing (km)", title = str_to_sentence(str_replace(i, "_", " ")))
    )
    
    print(i)
    
    # 0. Linear oxygen and temperature
    m0 <- sdmTMB(density ~ 0 + year_f + (1|month_f) + sal_sc + depth_sc + depth_sq + temp_sc + oxy_sc,
                 data = dd,
                 mesh = mesh,
                 #family = tweedie(link = "log"),
                 family = delta_gamma(type = "poisson-link"),
                 spatiotemporal = "iid",
                 spatial = "on",
                 time = "year")
    print("m0")
    sanity(m0)
    
    # 1. Linear oxygen and squared temperature
    m1 <- sdmTMB(density ~ 0 + year_f + (1|month_f) + sal_sc + depth_sc + depth_sq + temp_sc + temp_sq + oxy_sc,
                 data = dd,
                 mesh = mesh,
                 #family = tweedie(link = "log"),
                 family = delta_gamma(type = "poisson-link"),
                 spatiotemporal = "iid",
                 spatial = "on",
                 time = "year")
    print("m1")
    sanity(m1)
    
    # 2. Interaction between linear oxygen and temperature
    m2 <- sdmTMB(density ~ 0 + year_f + (1|month_f) + sal_sc + depth_sc + depth_sq + temp_sc * oxy_sc,
                  data = dd,
                  mesh = mesh,
                  #family = tweedie(link = "log"),
                  family = delta_gamma(type = "poisson-link"),
                  spatiotemporal = "iid",
                  spatial = "on",
                  time = "year")
    print("m2")
    sanity(m2)

    # 3. Breakpoint oxygen
    m3 <- sdmTMB(density ~ 0 + year_f + (1|month_f) + sal_sc + depth_sc + depth_sq + temp_sc + temp_sq + breakpt(oxy_sc),
                 data = dd,
                 mesh = mesh,
                 #family = tweedie(link = "log"),
                 family = delta_gamma(type = "poisson-link"),
                 spatiotemporal = "iid",
                 spatial = "on",
                 time = "year",
                 priors = sdmTMBpriors(
                   threshold_breakpt_slope = normal(0, 1),
                   threshold_breakpt_cut = normal(0, 1)))
    print("m3")
    sanity(m3)
    
    # 4. linear metabolic index
    m4 <- sdmTMB(density ~ 0 + year_f + (1|month_f) + sal_sc + depth_sc + depth_sq + phi_sc,
                 data = dd,
                 mesh = mesh,
                 #family = tweedie(link = "log"),
                 family = delta_gamma(type = "poisson-link"),
                 spatiotemporal = "iid",
                 spatial = "on",
                 time = "year")
    print("m4")
    sanity(m4)

    # 5. breakpoint metabolic index!
    m5 <- sdmTMB(density ~ 0 + year_f + (1|month_f) + sal_sc + depth_sc + depth_sq + breakpt(phi_sc),
                 data = dd,
                 mesh = mesh,
                 family = tweedie(link = "log"),
                 #family = delta_gamma(type = "poisson-link"),
                 spatiotemporal = "iid",
                 spatial = "on",
                 time = "year",
                 priors = sdmTMBpriors(
                   threshold_breakpt_slope = normal(0, 1),
                   threshold_breakpt_cut = normal(0, 1)))
                 
    print("m5")
    sanity(m5)
    
    # Save AIC
    AIC(m0, m1, m2, m3, m4, m5) |>
      tibble::rownames_to_column("model") |> 
      arrange(AIC)
    
    data_list_aic[[i]] <- AIC(m0, m1, m2, m3, m4, m5) |>
      tibble::rownames_to_column("model") |>
      mutate(group = i)
    
}

# Save aic as data frames
data_aic <- bind_rows(data_list_aic)

write_csv(data_aic, paste0(home, "/output/data_aic_01.csv"))
```


