---
title: "Fit models to biomass density and predict covariates"
author: "Max Lindmark"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document:
  df_print: paged
pdf_document: default
editor_options: 
  chunk_output_type: console
---

```{r setup, include = FALSE, cache=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 12,
  fig.asp = 0.618,
  fig.align ='center'
)
```

# TODO
1. Fix pred grid and standardization to account for different sizes!

# Intro
In this script, I take the collated stomach data set and calculate aggregates (feeding ratio, total weight of prey groups) and predictor variables for diet data, aggregate to get 1 stomach = 1 row per prey type (not prey individual). I also select only the columns I need for model fitting, join environmental covariates and cpue covariates for cod and flounder, and lastly saduria biomass densities.

## Load packages & source functions

```{r load libraries, message=FALSE, warning=TRUE}
# Load libraries, install if needed
library(tidyverse)
library(readxl)
library(tidylog)
library(RCurl)
library(RColorBrewer)
library(patchwork)
library(janitor)
library(forcats)
library(gapminder)
library(viridis)
library(ggridges)
library(raster)
library(icesDatras)
library(ggalluvial)
library(ggrepel)
library(ncdf4)
library(chron)
library(rnaturalearth)
library(rnaturalearthdata)
library(mapplots)
library(geosphere)
library(quantreg)
library(brms)
library(sdmTMB)
options(mc.cores = parallel::detectCores()) 

world <- ne_countries(scale = "medium", returnclass = "sf")

# Source code for map plots
# Source code for map plots
source("/Users/maxlindmark/Dropbox/Max work/R/spatial_metabolic_index/R/functions/map_plot.R")

plot_map_fc <- plot_map_fc + theme(legend.position = "bottom")

theme_set(theme_plot())

# Continuous colors
options(ggplot2.continuous.colour = "viridis")

# Discrete colors
scale_colour_discrete <- function(...) {
  scale_colour_brewer(palette = "Set2")
}

scale_fill_discrete <- function(...) {
  scale_fill_brewer(palette = "Set2")
}
```

## Read data

```{r read data}
# Load cache
# qwraps2::lazyload_cache_dir(path = "R/01_sdm_mi_cache/html")

pred_grid1 <- readr::read_csv("https://raw.githubusercontent.com/maxlindmark/cod_condition/master/data/for_analysis/pred_grid_(1_2).csv")
pred_grid2 <- readr::read_csv("https://raw.githubusercontent.com/maxlindmark/cod_condition/master/data/for_analysis/pred_grid_(2_2).csv")

pred_grid <- bind_rows(pred_grid1, pred_grid2)

pred_grid <- pred_grid %>%
  dplyr::select(-density_saduria, -biomass_spr, -biomass_her, -biomass_spr_sd, -biomass_her_sd, 
                -density_cod, -density_fle, -depth_rec, -temp_rec, -oxy_rec, -density_cod_rec, 
                -density_cod_sd, -density_fle_sd, -density_fle_rec, -density_saduria_rec, 
                -density_saduria_sd, -temp_sd, -oxy_sd) %>% 
  drop_na(oxy, temp)

cpue <- readr::read_csv("data/full_cpue_length_q1_q4.csv") %>%
  janitor::clean_names() %>% 
  rename(X = x,
         Y = y) %>% 
  filter(quarter %in% c(1, 4))

# Match pred grid else we cannot predict later
cpue <- cpue %>% filter(year %in% unique(pred_grid$year))
pred_grid

# Now scale variables in the cpue data by the mean and sd in the prediction grid
# pred_grid$log_depth <- log(pred_grid$depth)
# pred_grid$log_depth2 <- pred_grid$log_depth * pred_grid$log_depth
# mean_depth <- mean(pred_grid$log_depth)
# sd_depth <- sd(pred_grid$log_depth)
# mean_depth2 <- mean(pred_grid$log_depth2)
# sd_depth2 <- sd(pred_grid$log_depth2)
```

```{r}
# Calculating a metabolic index

# A_0 is the ratio of constant terms in the Arrhenius equations describing the rates of supply and demand
# B is individual body size (g)
# n is the allometric exponent (difference delta - epsilon)
# E0 is the difference in temperature dependence of supply and demand
# T is temperature in degrees Kelvin
# Tref is an arbitrarily chosen reference temperature (15°C)
# kb is Boltzmann’s constant (eV K−1).

# Oxygen unit conversion: 1 ml/l = 103/22.391 = 44.661 μmol/l

# Parameters from Deutsch et al 2015 Science Supp Fig. S2
A_conc = 9.7*10^-15
E_conc <- 0.72
n <- -0.21
Tref <- 273.15 + 10
kb <- 0.000086173324

pred_grid$T_K <- pred_grid$temp + 273.15
pred_grid$oxy2 <- (pred_grid$oxy * 10^3) / 22.391
pred_grid$phi <- A_conc*(2000^n)*(pred_grid$oxy2 / exp(-E_conc / (kb * pred_grid$T_K)))

# Scale variables in data w.r.t. prediction grid
mean_phi<- mean(pred_grid$phi)
sd_phi <- sd(pred_grid$phi)

mean_depth <- mean(pred_grid$depth)
sd_depth <- sd(pred_grid$depth)

mean_oxy <- mean(pred_grid$oxy)
sd_oxy <- sd(pred_grid$oxy)

mean_temp <- mean(pred_grid$temp)
sd_temp <- sd(pred_grid$temp)

# Calculate phi in data
# Metabolic index in the data
B_small_cod <- 0.01*median(filter(cpue, length_cm <= 30 & species == "cod")$length_cm)^3
B_large_cod <- 0.01*median(filter(cpue, length_cm > 30 & species == "cod")$length_cm)^3

# https://github.com/fate-spatialindicators/SDM_O2/blob/master/code/mi_functions.R
cpue <- cpue %>% 
  mutate(T_K = temp + 273.15,
         oxy2 = oxy * (10^3) / 22.391,
         size_cl = ifelse(length_cm > 30, "large", "small"),
         phi = ifelse(size_cl == "large", 
                      A_conc*(5000^n)*(oxy2 / exp(-E_conc / (kb * T_K))),
                      A_conc*(30^n)*(oxy2 / exp(-E_conc / (kb * T_K)))))

ggplot(cpue, aes(phi)) + geom_histogram() + facet_wrap(~size_cl, ncol = 1)

cpue <- cpue %>%
  mutate(
         # log_depth = log(depth),
         # log_depth2 = log_depth*log_depth,
         # depth_sc = (log_depth - mean_depth) / sd_depth,
         # depth_sc_sq = (log_depth2 - mean_depth2) / sd_depth2,
         depth_sc = (depth - mean_depth) / sd_depth,
         temp_sc = (temp - mean_temp) / sd_temp,
         oxy_sc = (oxy - mean_oxy) / sd_oxy,
         phi_sc = (phi - mean_phi) / sd_phi)
```

Use Essington data and fit mixed model, i.e., do not use Deutsch cod values straight off. Below I work with concentration instead of pressure, for simplicity, but will change later 

```{r}
# https://github.com/fate-spatialindicators/SDM_O2/blob/master/code/estimate_mi_parameters.R
# Read data:
library(rMR)

kb <- 0.000086173324

po2 <- read.csv("data/essington_pO2_extended.csv", sep = ";")

po2$po2 <- NA

spc_list <- unique(po2$spc)

c_to_K <- function(x) x + 273.15

for (i in 1:length(spc_list)) {
  
  spc_index <- which(po2$spc == spc_list[i])
  
  tmp_dat <- dplyr::filter(po2, spc == spc_list[i])
  
  tmp_dat$po2 <- DO.unit.convert(tmp_dat$lc50,
                                 DO.units.in = tmp_dat$units[1],
                                 DO.units.out = "PP",
                                 bar.press = 1,
                                 bar.units.in= "atm",
                                 temp.C = tmp_dat$temp,
                                 bar.units.out = "kpa",
                                 salinity = tmp_dat$salinity[1],
                                 salinity.units = "uS")
  
  po2$po2[spc_index] <- tmp_dat$po2
  
}

po2$spc <- as.factor(po2$spc)
nspc <- length(levels(po2$spc))

tref <- 15
po2$inv.temp <- (1 / kb)  * ( 1 / (po2$temp + 273.15) - 1 / (tref + 273.15))

fit <- glmmTMB(-log(po2) ~ diag(1+inv.temp|spc) + log(b), data = po2,
               family = gaussian(link = "identity"),
               se = TRUE
)

fit2 <- glmmTMB(-log(po2) ~ (1|spc) + log(b) + inv.temp, data = po2,
                family = gaussian(link = "identity"),
                se = TRUE
)

AICmat <- c(AIC(fit), AIC(fit2))
AICmat

fixef(fit2)
# Conditional model:
# (Intercept)       log(b)     inv.temp  
#    -0.05125     -0.28440      0.43290  
# ranef(fit2)
# $spc
#               (Intercept)
# AtlanticCod   -0.02088772

# Cod intercept:
cod_inter <- exp(fixef(fit2)$cond[1] + ranef(fit2)$cond$spc$`(Intercept)`[1])


spc <- levels(po2$spc)
n <- fixef(fit2)$cond[2]
Eo <- fixef(fit2)$cond[3]
Aospc_effects <- exp(fixef(fit2)$cond[1] + ranef(fit2)$cond$spc$`(Intercept)`)

po2$bnpo2 <- log(po2$b^n * po2$po2)



#species.names <- c("Atlantic cod", "Common eelpout", "Atlantic silverside", "N. swellfish", "Sablefish", "Sharpsnout sea bream", "Winter flounder")

for (i in 1:length(spc)) po2$Ao[po2$spc == spc[i]] <- Aospc_effects[i]
po2$Eo <- Eo

po2$bnpo2hat <- with(po2, -Eo * inv.temp - log(Ao) )

## TODO: Visualize fit for sablefish for range of temperatures observed in data
new.dat <- data_frame(spc = "sablefish",
                      b = 675,
                      inv.temp =(1/kb) * seq(1/c_to_K(13.6) - 1/c_to_K(15),
                                             1/c_to_K(2.98) - 1/c_to_K(15),
                                             length.out = 10)
)                        

fitted <- predict(fit, newdata = new.dat)
new.dat$po2 <- exp(-fitted)
new.dat$Ao <- Aospc_effects[7] # 5 in essingtons code
new.dat$Eo <- Eo
new.dat$bnpo2hat <- with(new.dat, -Eo * inv.temp - log(Ao))

# Plot
ggplot(data = filter(po2, !spc %in% c("flounder", "plaice")), aes(x = inv.temp, y = (bnpo2), col = spc)) + 
  geom_point(size = 3) +
  geom_line(size = 1, aes(y = bnpo2hat)) + 
  geom_line(data = new.dat, aes(x = inv.temp, y = bnpo2hat)) +
  scale_color_brewer(palette = "Accent", name = "") +
  #scale_color_brewer(palette = "Accent",name = "", labels = species_names) +
  #scale_color_brewer(palette = "Set1", name = "") +
  labs(x = bquote("Inverse Temperature"~(k[b]^{-1}~ (T^-1 - T[ref]^-1))), y = bquote('log'~(B^n~pO[2]))) +
  theme_bw() +
  theme(panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.border = element_blank()) +
  theme(axis.line = element_line(color = "black")) +
  theme(axis.text = element_text(size = 12)) +
  theme(axis.title= element_text(size = 14)) +
  theme(legend.text= element_text(size = 10)) +
  #theme(legend.position = c(0.8, 0.875))
  theme(legend.position = "bottom") + 
  theme(aspect.ratio = 1)


# Ok, now do the same but change unit! See how it compares for cod based on Deutsch values

# Finally, redo Essingtons plot for my 2 new species and in the new unit
for (i in 1:length(spc_list)) {
  
  spc_index <- which(po2$spc == spc_list[i])
  
  tmp_dat <- dplyr::filter(po2, spc == spc_list[i])
  
  tmp_dat$po2 <- DO.unit.convert(tmp_dat$lc50,
                                 DO.units.in = tmp_dat$units[1],
                                 DO.units.out = "mg/L", # Essington uses PP
                                 bar.press = 1,
                                 bar.units.in= "atm",
                                 temp.C = tmp_dat$temp,
                                 bar.units.out = "kpa",
                                 salinity = tmp_dat$salinity[1],
                                 salinity.units = "uS")
  
  po2$po2[spc_index] <- tmp_dat$po2
  
}

po2$spc <- as.factor(po2$spc)
nspc <- length(levels(po2$spc))

tref <- 15
po2$inv.temp <- (1 / kb)  * ( 1 / (po2$temp + 273.15) - 1 / (tref + 273.15))

fit2 <- glmmTMB(-log(po2) ~ (1|spc) + log(b) + inv.temp, data = po2,
                family = gaussian(link = "identity"),
                se = TRUE
)

# Now check the cod-specific intercept
fixef(fit2)
# Conditional model:
# (Intercept)       log(b)     inv.temp  
#      0.6922      -0.2855       0.2784  
ranef(fit2)
# $spc
#               (Intercept)
# AtlanticCod   -0.02420471

cod_inter2 <- exp(fixef(fit2)$cond[1] + ranef(fit2)$cond$spc$`(Intercept)`[1])
cod_inter2

# Parameters from Deutsch et al 2015 Science Supp Fig. S2
# I don't think this is centered... hence so small 
# https://github.com/fate-spatialindicators/SDM_O2/blob/master/code/mi_functions.R
A_conc = 9.7*10^-15

# What do I need to check here... I want to see if the concentration based metabolic index from Deutsch et al (2015) is the same when I use Essingtons equation for Deutsch parameters (the only difference should be temperature).

pred_grid$phi <- A_conc*(2000^n)*(pred_grid$oxy2 / exp(-E_conc / (kb * pred_grid$T_K)))

  
a <- 9.7*10^-15
E <- 0.72
Temp <- 283.15
b <- 1000
o2 <- (7 * 10^3) / 22.391 # 7 ml/L, common value. Convert to micro mol per liter (same as mg/L): https://www.unitsconverters.com/en/Micromol/L-To-Mg/Dl/Utu-6021-6020 
n <- -0.2
   
phi_d <- a*(b^n)*(o2/(exp(-E/(kb*Temp)))) 
phi_d

# Now do Essington's version... (but which constant? Cod estimate from their mixed model but concentration unit? Not the same then! Maybe I test essing and deutsch using pressure values... still, intercept is not the same lol)
  Ao <- 1.950318 # cod_inter2 # concentration intercept for cod
  kelvin = 273.15
  Temp2 <- 10
  
  inv.temp <- (1 / kb) * ( 1 / (Temp2 + kelvin) - 1 / (15 + kelvin)) # ref temp 10

  -E * inv_temp - A
  
phi_e <- (b^n)*Ao*o2 *exp(E * inv.temp)
phi_e
  
# This doesn't work. See if index in deutsch is the same as essing when using the pressure equation
a <- 3.1*10^-14
E <- 0.72
Temp <- 283.15
b <- 1000
o2 <- 103 # No idea what the partial pressure should be. According to this site: https://www.loligosystems.com/convert-oxygen-units using partial pressure at 50 m according to this site: https://bluerobotics.com/learn/pressure-depth-calculator/, 7 mg /l corresponds to 103 partial pressure
n <- -0.2
   
phi_d <- a*(b^n)*(o2/(exp(-E/(kb*Temp)))) 
phi_d

# Now do Essington:
Ao <- 0.9304069  # cod_inter
  
phi_e <- (b^n)*Ao*o2 *exp(E * inv.temp)
phi_e

# Ok, it's difficult... What if I take the Deutsch index for concentraiton straight as it is, then do the same for Essington parameters and just see how they compare?


# Parameters from Deutsch et al 2015 Science Supp Fig. S2
A_conc = 9.7*10^-15
E_conc <- 0.72
n <- -0.21
Tref <- 273.15 + 10
kb <- 0.000086173324

# Calculate phi in data
# Metabolic index in the data
B_small_cod <- 0.01*median(filter(cpue, length_cm <= 30 & species == "cod")$length_cm)^3
# small_cod_geom_mean <- cpue %>% 
#   filter(length_cm <= 30 & species == "cod") %>% 
#   filter(length_cm > 0) %>% 
#   summarise(geom_mean_weight = exp(mean(log(0.01*length_cm^3))))

B_large_cod <- 0.01*median(filter(cpue, length_cm > 30 & species == "cod")$length_cm)^3

# https://github.com/fate-spatialindicators/SDM_O2/blob/master/code/mi_functions.R
cpue <- cpue %>% 
  mutate(T_K = temp + 273.15,
         oxy2 = oxy * (10^3) / 22.391,
         size_cl = ifelse(length_cm > 30, "large", "small"),
         phi = ifelse(size_cl == "large", 
                      A_conc*(5000^n)*(oxy2 / exp(-E_conc / (kb * T_K))),
                      A_conc*(30^n)*(oxy2 / exp(-E_conc / (kb * T_K)))))

# Now do Essington (based on concentrations) and params from fit2
  Eo <- 0.2784
  #Eo <- 0.72
  Ao <-  0.1422288 # 1.950318 is Cod inter 2. This intercept = -log(Ao)
  #n <- -0.2855
  n <- -0.21
  kelvin = 273.15

cpue <- cpue %>% 
  mutate(inv.temp = (1 / kb) * ( 1 / (temp + kelvin) - 1 / (15 + kelvin)),
         phi_es = ifelse(size_cl == "large", 
                         5000^n*Ao*oxy2*exp(Eo * inv.temp),
                         1000^n*Ao*oxy2*exp(Eo * inv.temp)))
  
ggplot(cpue) +
  facet_wrap(~size_cl, ncol = 1, scales = "free_y") +
  geom_histogram(aes(phi, fill = "deutsch_conc"), color = NA, alpha = 0.5) +
  geom_histogram(aes(phi_es, fill = "essing_conc"), color = NA, alpha = 0.5)
# This plot looks better if I divide essington intercept by 2.

# Summary:
# Ok this looks reasonable
# What i need to understand now is the following:
# According to Deutsch (2015) it doesn't really matter if I work with pressure or concentration. The parameters change, but they also cancel each other out sort of.
# I want to use Essington's way of estimating things. He uses pressure though.
# Is it OK to simply switch units of any variable related to oxygen (and which are those?)
# How to find out: does the equation change? NO... Check the conversion to mg/L in the function: anything I need to sort out there?
# Ultimate test: calculate some values in essingtons data, then refit the model with concentraton to get new pO2_crit values, convert data to concentration again then then calculate. Basically, do Deeutsch appendix plot with essing data.

```

## Summarise data

```{r aggregate cpue per length to cpue per haul per size-class}
# Summaries density by length group per haul
cod_cpue <- cpue %>%
  filter(species == "cod") %>%
  mutate(size_group = ifelse(length_cm >= 30, "large", "small")) %>% 
  group_by(haul_id, size_group) %>% 
  summarise(density = sum(density)) %>% 
  ungroup()

  cod_cpue_small <- cod_cpue %>%
    filter(size_group == "small") %>%
    dplyr::select(-size_group) %>% 
    rename(density_cod_small = density)

  cod_cpue_large <- cod_cpue %>%
    filter(size_group == "large") %>%
    dplyr::select(-size_group) %>% 
    rename(density_cod_large = density)

fle_cpue <- cpue %>%
  filter(species == "flounder") %>%
  mutate(size_group = ifelse(length_cm >= 22, "large", "small")) %>% 
  group_by(haul_id, size_group) %>% 
  summarise(density = sum(density)) %>% 
  ungroup()

  fle_cpue_small <- fle_cpue %>%
    filter(size_group == "small") %>%
    dplyr::select(-size_group) %>% 
    rename(density_fle_small = density)

  fle_cpue_large <- fle_cpue %>%
    filter(size_group == "large") %>%
    dplyr::select(-size_group) %>% 
    rename(density_fle_large = density)


# Check 0 catches
cod_cpue_small %>%
  group_by(haul_id) %>%
  summarise(haul_dens = sum(density_cod_small)) %>%
  ungroup() %>%
  filter(!haul_dens == 0)

cod_cpue_large %>%
  group_by(haul_id) %>%
  summarise(haul_dens = sum(density_cod_large)) %>%
  ungroup() %>%
  filter(!haul_dens == 0)

fle_cpue_small %>%
  group_by(haul_id) %>%
  summarise(haul_dens = sum(density_fle_small)) %>%
  ungroup() %>%
  filter(!haul_dens == 0)

fle_cpue_large %>%
  group_by(haul_id) %>%
  summarise(haul_dens = sum(density_fle_large)) %>%
  ungroup() %>%
  filter(!haul_dens == 0)
```

```{r fit models and predict on stomach data}
cod_cpue_small <- left_join(cod_cpue_small, 
                            cpue %>%
                              filter(species == "cod") %>% 
                              distinct(haul_id, .keep_all = TRUE) %>%
                              dplyr::select(haul_id, year, quarter, X, Y, oxy, temp, depth,
                                             oxy_sc, temp_sc, depth_sc, phi, phi_sc)) %>% 
  drop_na(phi)

cod_cpue_large <- left_join(cod_cpue_large,
                            cpue %>%
                              filter(species == "cod") %>% 
                              distinct(haul_id, .keep_all = TRUE) %>%
                              dplyr::select(haul_id, year, quarter, X, Y, oxy, temp, depth,
                                             oxy_sc, temp_sc, depth_sc, phi, phi_sc)) %>% 
  drop_na(phi)

fle_cpue_small <- left_join(fle_cpue_small, 
                            cpue %>%
                              filter(species == "flounder") %>% 
                              distinct(haul_id, .keep_all = TRUE) %>%
                              dplyr::select(haul_id, year, quarter, X, Y, oxy, temp, depth,
                                             oxy_sc, temp_sc, depth_sc))

fle_cpue_large <- left_join(fle_cpue_large,
                            cpue %>%
                              filter(species == "flounder") %>% 
                              distinct(haul_id, .keep_all = TRUE) %>%
                              dplyr::select(haul_id, year, quarter, X, Y, oxy, temp, depth,
                                            oxy_sc, temp_sc, depth_sc))
```

## Fit models

```{r make meshes}
# Small cod
spde_small_cod <- make_mesh(cod_cpue_small, xy_cols = c("X", "Y"),
                            n_knots = 150, 
                            type = "kmeans", seed = 42)


# Large cod
spde_large_cod <- make_mesh(cod_cpue_large, xy_cols = c("X", "Y"),
                            n_knots = 150, 
                            type = "kmeans", seed = 42)

# Small flounder
spde_small_fle <- make_mesh(fle_cpue_small, xy_cols = c("X", "Y"),
                            n_knots = 150, 
                            type = "kmeans", seed = 42)


# Large flounder
spde_large_fle <- make_mesh(fle_cpue_large, xy_cols = c("X", "Y"),
                            n_knots = 150, 
                            type = "kmeans", seed = 42)

plot(spde_small_cod)
plot(spde_large_cod)
plot(spde_small_fle)
plot(spde_large_fle)
```

Small cod

```{r fit small cod model, cache=TRUE}
# Small cod model
mcod_s <- sdmTMB(density_cod_small ~ 0 + as.factor(quarter) + as.factor(year) + s(depth_sc) + oxy_sc + temp_sc,
                 data = cod_cpue_small,
                 mesh = spde_small_cod,
                 family = tweedie(link = "log"),
                 spatiotemporal = "off",
                 spatial = "on",
                 time = "year",
                 reml = FALSE)

mcod_s_v2 <- run_extra_optimization(mcod_s, nlminb_loops = 1, newton_loops = 1)
```

```{r}
sanity(mcod_s_v2)
```

```{r fit small cod model1, cache=TRUE}
# Small cod model
mcod_s1 <- sdmTMB(density_cod_small ~ 0 + as.factor(quarter) + as.factor(year) + s(depth_sc) + breakpt(oxy_sc) + temp_sc,
                 data = cod_cpue_small,
                 mesh = spde_small_cod,
                 family = tweedie(link = "log"),
                 spatiotemporal = "off",
                 spatial = "on",
                 time = "year",
                 reml = FALSE)

mcod_s1_v2 <- run_extra_optimization(mcod_s1, nlminb_loops = 1, newton_loops = 1)
```

```{r}
sanity(mcod_s1_v2)
```

```{r fit small cod model2, cache=TRUE}
# Small cod model
mcod_s2 <- sdmTMB(density_cod_small ~ 0 + as.factor(quarter) + as.factor(year) + s(depth_sc) + oxy_sc * temp_sc,
                 data = cod_cpue_small,
                 mesh = spde_small_cod,
                 family = tweedie(link = "log"),
                 spatiotemporal = "off",
                 spatial = "on",
                 time = "year",
                 reml = FALSE)

mcod_s2_v2 <- run_extra_optimization(mcod_s2, nlminb_loops = 1, newton_loops = 1)
```

```{r}
sanity(mcod_s2_v2)
```

```{r fit small cod model3, cache=TRUE}
# Small cod model
mcod_s3 <- sdmTMB(density_cod_small ~ 0 + as.factor(quarter) + as.factor(year) + s(depth_sc) + breakpt(phi_sc),
                 data = cod_cpue_small,
                 mesh = spde_small_cod,
                 family = tweedie(link = "log"),
                 spatiotemporal = "off",
                 spatial = "on",
                 time = "year",
                 reml = FALSE)

mcod_s3_v2 <- run_extra_optimization(mcod_s3, nlminb_loops = 1, newton_loops = 1)
```

```{r}
sanity(mcod_s3_v2)
```

```{r fit small cod model4, cache=TRUE}
# Small cod model
mcod_s4 <- sdmTMB(density_cod_small ~ 0 + as.factor(quarter) + as.factor(year) + s(depth_sc) + s(phi_sc),
                 data = cod_cpue_small,
                 mesh = spde_small_cod,
                 family = tweedie(link = "log"),
                 spatiotemporal = "off",
                 spatial = "on",
                 time = "year",
                 reml = FALSE)

mcod_s4_v2 <- run_extra_optimization(mcod_s4, nlminb_loops = 1, newton_loops = 1)
```

```{r}
sanity(mcod_s4_v2)
```


```{r}
#mcmc_res_sc <- residuals(mcod_s_v2, type = "mle-mcmc", mcmc_iter = 201, mcmc_warmup = 200)
#qqnorm(mcmc_res_sc);qqline(mcmc_res_sc)
```

Large cod

```{r fit large cod model, cache=TRUE}
# Large cod model
mcod_l <- sdmTMB(density_cod_large ~ 0 + as.factor(quarter) + as.factor(year) + s(depth_sc) + oxy_sc + temp_sc,
                 data = cod_cpue_large,
                 mesh = spde_large_cod,
                 family = tweedie(link = "log"),
                 spatiotemporal = "off",
                 spatial = "on",
                 time = "year",
                 reml = FALSE)

mcod_l_v2 <- run_extra_optimization(mcod_l, nlminb_loops = 1, newton_loops = 1)
```

```{r}
sanity(mcod_l_v2)
```

```{r fit large cod model1, cache=TRUE}
# Large cod model
mcod_l1 <- sdmTMB(density_cod_large ~ 0 + as.factor(quarter) + as.factor(year) + s(depth_sc) + breakpt(oxy_sc) + temp_sc,
                 data = cod_cpue_large,
                 mesh = spde_large_cod,
                 family = tweedie(link = "log"),
                 spatiotemporal = "off",
                 spatial = "on",
                 time = "year",
                 reml = FALSE)

mcod_l1_v2 <- run_extra_optimization(mcod_l1, nlminb_loops = 1, newton_loops = 1)
```

```{r}
sanity(mcod_l1_v2)
```

```{r fit large cod model2, cache=TRUE}
# Large cod model
mcod_l2 <- sdmTMB(density_cod_large ~ 0 + as.factor(quarter) + as.factor(year) + s(depth_sc) + oxy_sc*temp_sc,
                 data = cod_cpue_large,
                 mesh = spde_large_cod,
                 family = tweedie(link = "log"),
                 spatiotemporal = "off",
                 spatial = "on",
                 time = "year",
                 reml = FALSE)

mcod_l2_v2 <- run_extra_optimization(mcod_l2, nlminb_loops = 1, newton_loops = 1)
```

```{r}
sanity(mcod_l2_v2)
```

```{r fit large cod model3, cache=TRUE}
# Large cod model
mcod_l3 <- sdmTMB(density_cod_large ~ 0 + as.factor(quarter) + as.factor(year) + s(depth_sc) + breakpt(phi_sc),
                 data = cod_cpue_large,
                 mesh = spde_large_cod,
                 family = tweedie(link = "log"),
                 spatiotemporal = "off",
                 spatial = "on",
                 time = "year",
                 reml = FALSE)

mcod_l3_v2 <- run_extra_optimization(mcod_l3, nlminb_loops = 1, newton_loops = 1)
```

```{r}
sanity(mcod_l3_v2)
```

```{r fit large cod model4, cache=TRUE}
# Large cod model
mcod_l4 <- sdmTMB(density_cod_large ~ 0 + as.factor(quarter) + as.factor(year) + s(depth_sc) + s(phi_sc),
                 data = cod_cpue_large,
                 mesh = spde_large_cod,
                 family = tweedie(link = "log"),
                 spatiotemporal = "off",
                 spatial = "on",
                 time = "year",
                 reml = FALSE)

mcod_l4_v2 <- run_extra_optimization(mcod_l4, nlminb_loops = 1, newton_loops = 1)
```

```{r}
sanity(mcod_l4_v2)
```

```{r}
#mcmc_res_lc <- residuals(mcod_l_v2, type = "mle-mcmc", mcmc_iter = 201, mcmc_warmup = 200)
#qqnorm(mcmc_res_lc);qqline(mcmc_res_lc)
```

Small flounder

```{r fit small flounder model, cache=TRUE}
# Small flounder model
# mfle_s <- sdmTMB(density_fle_small ~ 0 + as.factor(quarter) + as.factor(year) + s(depth_sc) +
#                    breakpt(oxy_sc) + temp_sc,
#                  data = fle_cpue_small,
#                  mesh = spde_small_fle,
#                  family = tweedie(link = "log"),
#                  spatiotemporal = "off",
#                  spatial = "on",
#                  time = "year",
#                  reml = FALSE)
# 
# mfle_s_v2 <- run_extra_optimization(mfle_s, nlminb_loops = 1, newton_loops = 1)
```

```{r fit small flounder model2, cache=TRUE}
# Small flounder model
# mfle_s2 <- sdmTMB(density_fle_small ~ 0 + as.factor(quarter) + as.factor(year) + s(depth_sc) +
#                    breakpt(oxy_sc) + temp_sc,
#                  data = fle_cpue_small,
#                  mesh = spde_small_fle,
#                  family = tweedie(link = "log"),
#                  spatiotemporal = "off",
#                  spatial = "on",
#                  time = "year",
#                  reml = FALSE)
# 
# mfle_s2_v2 <- run_extra_optimization(mfle_s2, nlminb_loops = 1, newton_loops = 1)
```

```{r}
#mcmc_res_sf <- residuals(mfle_s_v2, type = "mle-mcmc", mcmc_iter = 201, mcmc_warmup = 200)
#qqnorm(mcmc_res_sf);qqline(mcmc_res_sf)
```

Large flounder

```{r fit large flounder model, cache=TRUE}
# Large flounder model
# mfle_l <- sdmTMB(density_fle_large ~ 0 + as.factor(quarter) + as.factor(year) + s(depth_sc) +
#                    breakpt(oxy_sc) + temp_sc,
#                  data = fle_cpue_large,
#                  mesh = spde_large_fle,
#                  family = tweedie(link = "log"),
#                  spatiotemporal = "off",
#                  spatial = "on",
#                  time = "year",
#                  reml = FALSE)
# 
# mfle_l_v2 <- run_extra_optimization(mfle_l, nlminb_loops = 1, newton_loops = 1)
```

```{r fit large flounder model2, cache=TRUE}
# Large flounder model
# mfle_l2 <- sdmTMB(density_fle_large ~ 0 + as.factor(quarter) + as.factor(year) + s(depth_sc) +
#                    oxy_sc*temp_sc,
#                  data = fle_cpue_large,
#                  mesh = spde_large_fle,
#                  family = tweedie(link = "log"),
#                  spatiotemporal = "off",
#                  spatial = "on",
#                  time = "year",
#                  reml = FALSE)
# 
# mfle_l2_v2 <- run_extra_optimization(mfle_l2, nlminb_loops = 1, newton_loops = 1)
```

```{r}
#mcmc_res_lf <- residuals(mfle_l_v2, type = "mle-mcmc", mcmc_iter = 201, mcmc_warmup = 200)
#qqnorm(mcmc_res_lf);qqline(mcmc_res_lf)
```

## Results
### AIC
```{r aic}
AIC(mcod_s_v2, mcod_s1_v2, mcod_s2_v2, mcod_s3_v2, mcod_s4_v2)
AIC(mcod_l_v2, mcod_l1_v2, mcod_l2_v2, mcod_l3_v2, mcod_l4_v2)
```

For both small and large cod, the smooth phi model is best

```{r}
ggplot(cod_cpue_large, aes(phi, density_cod_large)) + geom_point() + stat_smooth()
ggplot(cod_cpue_small, aes(phi, density_cod_small)) + geom_point() + stat_smooth()

tidy(mcod_l4_v2)
tidy(mcod_s4_v2)

large_pars <- tidy(mcod_l2_v2,
                effects = "fixed", conf.int = TRUE) %>%
  filter(!grepl('year', term)) %>%
  filter(!grepl('quarter', term)) %>% 
  mutate(size_cl = "Adult")

small_pars <- tidy(mcod_s2_v2,
                effects = "fixed", conf.int = TRUE) %>%
  filter(!grepl('year', term)) %>%
  filter(!grepl('quarter', term)) %>% 
  mutate(size_cl = "Juvenile")

coefs <- bind_rows(large_pars, small_pars)

coefs <- coefs %>% 
    mutate(term2 = term,
           term2 = ifelse(term == "temp_sc", "Temperature", term2),
           term2 = ifelse(term == "oxy_sc", "Oxygen", term2),
           term2 = ifelse(term == "depth_sc", "Depth", term2),
           term2 = ifelse(term == "oxy_sc:temp_sc", "Temperature x Oxygen", term2)
           )

coefs <- coefs %>% 
    mutate(term2 = term,
           term2 = ifelse(term == "temp_sc", "Temperature", term2),
           term2 = ifelse(term == "oxy_sc", "Oxygen", term2),
           term2 = ifelse(term == "depth_sc", "Depth", term2),
           term2 = ifelse(term == "oxy_sc:temp_sc", "Temperature x Oxygen", term2)
           )

# Coefficients
ggplot(coefs, aes(term2, estimate, color = size_cl)) +
  geom_hline(yintercept = 0, linetype = 2, color = "gray50", size = 0.5) +
  geom_point(size = 3, position = position_dodge(width = 0.2)) +
  geom_errorbar(aes(ymin = conf.low, ymax = conf.high), width = 0, size = 1,
                position = position_dodge(width = 0.2)) +
  labs(x = "Predictor", y = "Standardized coefficient") +
  theme(legend.position = "right",
        axis.text.x = element_text(angle = 90)) +
  NULL 
```

### 2. Marginal effects plot

Weird that it's unimodal and declines at high phi's. Maybe super coastal?

```{r margin phi}
# Marginal
nd <- data.frame(
  phi_sc = seq(min(cod_cpue_small$phi_sc) + 0.5,
               max(cod_cpue_small$phi_sc) - 0.2, length.out = 100), 
  year = 2015L,
  quarter = 1,
  oxy_sc = mean(cod_cpue_small$oxy_sc),
  temp_sc = mean(cod_cpue_small$temp_sc),
  depth_sc = mean(cod_cpue_small$depth_sc)
)

p_adu <- predict(mcod_l4_v2, newdata = nd, se_fit = TRUE, re_form = NA, xy_cols = c("X", "Y"))
p_juv <- predict(mcod_s4_v2, newdata = nd, se_fit = TRUE, re_form = NA, xy_cols = c("X", "Y"))

p <- bind_rows(p_adu %>% mutate(size_cl = "Adult"),
               p_juv %>% mutate(size_cl = "Juvenile")) %>% 
  mutate(phi = (phi_sc * sd_phi) + mean_phi)

ggplot(p, aes(phi, exp(est), color = size_cl, fill = size_cl, 
              ymin = exp(est - 1.96 * est_se), 
              ymax = exp(est + 1.96 * est_se))) +
  geom_line() +
  geom_ribbon(alpha = 0.4, color = NA) + 
  geom_vline(xintercept = 1, linetype = 2, color = "grey30") +
  coord_cartesian(expand = 0) + 
  scale_x_continuous(breaks = seq(-2, 40, by = 2))
```

3. Predictions on maps

```{r map pred}

pred_grid

plot_map_fc + 
  geom_raster(data = pred_grid, aes(X*1000, Y*1000, fill = phi)) + 
  facet_wrap(~year) + 
  scale_fill_viridis()
```

4. weighted phi by size group and year

```{r weighted means}
pred_grid <- pred_grid %>%
  mutate(quarter = factor(4),
         depth_sc = (depth - mean(depth)) / sd(depth),
         phi_sc = (phi - mean(phi)) / sd(phi))

large_pred <- predict(mcod_l4_v2, newdata = pred_grid)
small_pred <- predict(mcod_s4_v2, newdata = pred_grid)

# Small cod
wm_phi_small <- small_pred %>%
  group_by(year) %>%
  summarise("Median" = hutils::weighted_quantile(v = phi, w = exp(est), p = c(0.5)),
            "1st quartile" = hutils::weighted_quantile(v = phi, w = exp(est), p = c(0.25)),
            "3rd quartile" = hutils::weighted_quantile(v = phi, w = exp(est), p = c(0.75))) %>% 
  ungroup() %>% 
  mutate(group = "Juvenile") 

# Large cod
wm_phi_large <- large_pred %>%
  group_by(year) %>%
  summarise("Median" = hutils::weighted_quantile(v = phi, w = exp(est), p = c(0.5)),
            "1st quartile" = hutils::weighted_quantile(v = phi, w = exp(est), p = c(0.25)),
            "3rd quartile" = hutils::weighted_quantile(v = phi, w = exp(est), p = c(0.75))) %>% 
  ungroup() %>% 
  mutate(group = "Adult")

env_phi <- pred_grid %>%
  group_by(year) %>%
  summarise("Median" = quantile(phi, p = c(0.5)),
            "1st quartile" = quantile(phi, p = c(0.25)),
            "3rd quartile" = quantile(phi, p = c(0.75))) %>% 
  ungroup() %>% 
  mutate(group = "Environment") 
  
phi_series <- bind_rows(wm_phi_small, wm_phi_large, env_phi)

ggplot(phi_series, aes(year, Median, color = group, fill = group)) +
  geom_line() +
  geom_ribbon(aes(ymin = `1st quartile`, ymax = `3rd quartile`), alpha = 0.2, color = NA) +
  facet_wrap(~group) +
  scale_x_continuous(breaks = scales::pretty_breaks(n = 6)) +
  guides(fill = "none", color = "none") +
  labs(y = "phi", x = "Year") +
  theme_plot() +
  theme(axis.text.x = element_text(angle = 90)) +
  NULL

ggplot(phi_series, aes(year, Median, color = group)) +
  geom_line() +
  scale_x_continuous(breaks = scales::pretty_breaks(n = 6)) +
  labs(y = "phi", x = "Year", color = "") +
  theme_plot() +
  theme(axis.text.x = element_text(angle = 90)) +
  NULL
```

