---
title: "Plot conditional effects and grid predictions from 02"
author: "Max Lindmark"
date: today
date-format: iso
toc: true
format: 
  html:
    page-layout: full
    embed-resources: true
knitr: 
  opts_chunk:
    fig.align: center
    out-width: 100%
editor: source
---

## Intro

Plot conditional effects from predictions in script "02-fit-sdms-conditional.qmd"

## Load packages & source functions

```{r load libraries}
#| message: false
#| warning: false

# Load libraries, install if needed
library(tidyverse)
library(readxl)
library(tidylog)
library(RCurl)
library(devtools)
library(patchwork)
library(viridis)
library(RColorBrewer)
library(here)
library(ggnewscale)
library(ggh4x)
library(egg)
library(ggstats)
library(concaveman)

devtools::source_url("https://raw.githubusercontent.com/maxlindmark/spatial-metabolic-index/main/R/functions/map-plot.R")
options(ggplot2.continuous.colour = "viridis")

pal_temp <- brewer.pal(n = 10, name = "Paired")[c(2, 6)]
pal_oxy <- brewer.pal(n = 10, name = "Paired")[c(8, 4)]

# Set path
home <- here::here()
```

## Read predictions

```{r}
#| warning: false
#| message: false

preds <- read_csv(paste0(home, "/output/conditional_02_sdms.csv")) |>
  filter(oxy > 3 & oxy < 9) |>
  filter(temp > 2 & temp < 14) |>
  separate(group, c("species", "life_stage"), sep = "_", remove = FALSE) |>
  mutate(
    species = str_to_title(species),
    life_stage = str_to_title(life_stage),
    group2 = paste(species, life_stage),
    group2 = factor(group2,
      levels = c(
        "Cod Adult", "Cod Juvenile", "Flounder Adult", "Flounder Juvenile",
        "Plaice Adult", "Plaice Juvenile"
      )
    )
  ) |>
  mutate(
    est_sc = exp(est) / max(exp(est)),
    .by = group
  ) # for plotting all together in heatmap

grid_preds <- read_csv(paste0(home, "/output/pred_grids_02_sdms.csv")) |>
  separate(group, c("species", "life_stage"), sep = "_", remove = FALSE) |>
  mutate(
    species = str_to_title(species),
    life_stage = str_to_title(life_stage),
    group2 = paste(species, life_stage),
    group2 = factor(group2,
      levels = c(
        "Cod Adult", "Cod Juvenile", "Flounder Adult", "Flounder Juvenile",
        "Plaice Adult", "Plaice Juvenile"
      )
    )
  ) |>
  mutate(est = exp(est))

cogs <- read_csv(paste0(home, "/output/cogs_02_sdms.csv")) |>
  separate(group, c("species", "life_stage"), sep = "_", remove = FALSE) |>
  mutate(
    species = str_to_title(species),
    life_stage = str_to_title(life_stage)
  )

eaos <- read_csv(paste0(home, "/output/eaos_02_sdms.csv")) |>
  separate(group, c("species", "life_stage"), sep = "_", remove = FALSE) |>
  mutate(
    species = str_to_title(species),
    life_stage = str_to_title(life_stage),
    group2 = paste(species, life_stage),
    group2 = factor(group2,
      levels = c(
        "Cod Adult", "Cod Juvenile", "Flounder Adult", "Flounder Juvenile",
        "Plaice Adult", "Plaice Juvenile"
      )
    )
  )

index <- read_csv(paste0(home, "/output/index_02_sdms.csv")) |>
  separate(group, c("species", "life_stage"), sep = "_", remove = FALSE) |>
  mutate(
    species = str_to_title(species),
    life_stage = str_to_title(life_stage),
    group2 = paste(species, life_stage),
    group2 = factor(group2,
      levels = c(
        "Cod Adult", "Cod Juvenile", "Flounder Adult", "Flounder Juvenile",
        "Plaice Adult", "Plaice Juvenile"
      )
    )
  )

scenario_pred <- read_csv(paste0(home, "/output/scenario_preds_02_sdms.csv")) |>
  separate(group, c("species", "life_stage"), sep = "_", remove = FALSE) |>
  mutate(
    species = str_to_title(species),
    life_stage = str_to_title(life_stage),
    group2 = paste(species, life_stage),
    group2 = factor(group2,
      levels = c(
        "Cod Adult", "Cod Juvenile", "Flounder Adult", "Flounder Juvenile",
        "Plaice Adult", "Plaice Juvenile"
      )
    )
  )
```

## Calculate weighted quantiles

```{r}
#| warning: false
#| message: false

wq <- grid_preds |>
  summarise(
    "Median_oxy" = hutils::weighted_quantile(v = oxy, w = est, p = 0.5),
    "10_oxy" = hutils::weighted_quantile(v = oxy, w = est, p = 0.1),
    "25_oxy" = hutils::weighted_quantile(v = oxy, w = est, p = 0.25),
    "75_oxy" = hutils::weighted_quantile(v = oxy, w = est, p = 0.75),
    "90_oxy" = hutils::weighted_quantile(v = oxy, w = est, p = 0.9),
    "Env_oxy" = median(oxy),
    "Median_temp" = hutils::weighted_quantile(v = temp, w = est, p = 0.5),
    "10_temp" = hutils::weighted_quantile(v = temp, w = est, p = 0.1),
    "25_temp" = hutils::weighted_quantile(v = temp, w = est, p = 0.25),
    "75_temp" = hutils::weighted_quantile(v = temp, w = est, p = 0.75),
    "90_temp" = hutils::weighted_quantile(v = temp, w = est, p = 0.9),
    "Env_temp" = median(temp),
    "Median_sal" = hutils::weighted_quantile(v = sal, w = est, p = 0.5),
    "10_sal" = hutils::weighted_quantile(v = sal, w = est, p = 0.1),
    "25_sal" = hutils::weighted_quantile(v = sal, w = est, p = 0.25),
    "75_sal" = hutils::weighted_quantile(v = sal, w = est, p = 0.75),
    "90_sal" = hutils::weighted_quantile(v = sal, w = est, p = 0.9),
    "Env_sal" = median(sal),
    .by = c(group2, quarter)
  ) |>
  separate(group2, into = c("species", "life_stage"), remove = FALSE)

wq_annual <- grid_preds |>
  summarise(
    "Median_oxy" = hutils::weighted_quantile(v = oxy, w = est, p = 0.5),
    "1st_oxy" = hutils::weighted_quantile(v = oxy, w = est, p = 0.25),
    "3rd_oxy" = hutils::weighted_quantile(v = oxy, w = est, p = 0.75),
    "Env_oxy" = median(oxy),
    "Median_temp" = hutils::weighted_quantile(v = temp, w = est, p = 0.5),
    "1st_temp" = hutils::weighted_quantile(v = temp, w = est, p = 0.25),
    "3rd_temp" = hutils::weighted_quantile(v = temp, w = est, p = 0.75),
    "Env_temp" = median(temp),
    "Median_sal" = hutils::weighted_quantile(v = sal, w = est, p = 0.5),
    "1st_sal" = hutils::weighted_quantile(v = sal, w = est, p = 0.25),
    "3rd_sal" = hutils::weighted_quantile(v = sal, w = est, p = 0.75),
    "Env_sal" = median(sal),
    .by = c(group2, year, quarter)
  ) |>
  separate(group2, into = c("species", "life_stage"), remove = FALSE)
```

## Plot conditional effects

```{r}
#| warning: false
#| message: false

p1 <- ggplot(preds, aes(temp, oxy, fill = est_sc)) +
  facet_grid(life_stage ~ species) +
  geom_raster(alpha = 0.95) +
  geom_contour(aes(temp, oxy, z = est_sc),
    color = "grey95",
    bins = 10, linewidth = 0.3, alpha = 0.5, inherit.aes = FALSE
  ) +
  geom_point(data = wq, aes(Median_temp, Median_oxy,
    color = as.factor(quarter)
  ), inherit.aes = FALSE) +
  geom_errorbar(
    data = wq, aes(Median_temp, Median_oxy,
      xmin = `25_temp`, xmax = `75_temp`,
      color = as.factor(quarter)
    ), inherit.aes = FALSE,
    alpha = 0.7, width = 0
  ) +
  geom_errorbar(
    data = wq, aes(Median_temp, Median_oxy,
      ymin = `25_oxy`, ymax = `75_oxy`,
      color = as.factor(quarter)
    ), inherit.aes = FALSE,
    alpha = 0.7, width = 0
  ) +
  scale_fill_viridis() +
  scale_color_manual(values = c("grey90", "tomato3")) +
  labs(
    y = "Oxygen (ml/L)",
    x = "Temperature (°C)",
    color = "Quarter",
    fill = expression("Scaled biomass density (kg/km"^2 * ")")
  ) +
  guides(
    color = guide_legend(title.position = "top", title.hjust = 0.5),
    fill = guide_colorbar(title.position = "top", title.hjust = 0.5)
  ) +
  theme(
    legend.position = "bottom",
    legend.key.width = unit(1.2, "cm"),
    legend.key.height = unit(0.2, "cm"),
    legend.title.position = "top"
  ) +
  coord_cartesian(expand = 0) +
  NULL

tag_facet(p1, fontface = 1, col = "gray95", size = 3)

ggsave(paste0(home, "/figures/temp_oxy_conditional.pdf"), width = 18, height = 13, units = "cm")
```

```{r}
#| warning: false
#| message: false

# Oxygen
wi_median_oxy <- preds |>
  left_join(wq |>
    filter(quarter == 4) |>
    dplyr::select(group2, Median_oxy), by = "group2") |>
  distinct(group2, Median_oxy) |>
  separate(group2, into = c("species", "life_stage"), remove = FALSE)

wi_median_temp <- preds |>
  left_join(wq |>
    filter(quarter == 4) |>
    dplyr::select(group2, Median_temp), by = "group2") |>
  distinct(group2, Median_temp) |>
  separate(group2, into = c("species", "life_stage"), remove = FALSE)

ggplot() +
  facet_grid2(life_stage ~ species, independent = "y", scales = "free_y") +
  geom_vline(
    data = wi_median_oxy, aes(xintercept = Median_oxy),
    linetype = 2, alpha = 0.5, linewidth = 0.4
  ) +
  geom_line(data = preds, aes(oxy, est, color = temp, group = temp)) +
  scale_color_viridis(option = "plasma") +
  labs(
    x = "Oxygen (ml/L)",
    y = expression(paste("Log biomass density (kg/km"^2, ")")),
    color = "Temperature (°C)"
  ) +
  guides(color = guide_colorbar(title.position = "top", title.hjust = 0.5)) +
  theme(
    legend.position = "bottom",
    legend.key.width = unit(1.2, "cm"),
    legend.key.height = unit(0.2, "cm"),
    legend.title.position = "top"
  ) +
  coord_cartesian(expand = 0)

ggsave(paste0(home, "/figures/supp/oxy_conditional.pdf"), width = 19, height = 10, units = "cm")

# Temperature
ggplot() +
  facet_grid2(life_stage ~ species, independent = "y", scales = "free_y") +
  geom_vline(
    data = wi_median_temp, aes(xintercept = Median_temp),
    linetype = 2, alpha = 0.5, linewidth = 0.4
  ) +
  geom_line(data = preds, aes(temp, est, color = oxy, group = oxy)) +
  labs(
    x = "Temperature (°C)",
    y = expression(paste("Log biomass density (kg/km"^2, ")")),
    color = "Oxygen (ml/L)"
  ) +
  guides(color = guide_colorbar(title.position = "top", title.hjust = 0.5)) +
  theme(
    legend.position = "bottom",
    legend.key.width = unit(1.2, "cm"),
    legend.key.height = unit(0.2, "cm"),
    legend.title.position = "top"
  ) +
  coord_cartesian(expand = 0)

ggsave(paste0(home, "/figures/supp/temp_conditional.pdf"), width = 19, height = 10, units = "cm")
```

## Plot weighted quantiles over time

```{r}
#| warning: false
#| message: false

t <- wq_annual |>
  pivot_longer(c(
    "Median_oxy", `1st_oxy`, `3rd_oxy`, "Env_oxy",
    "Median_temp", `1st_temp`, `3rd_temp`, "Env_temp",
    "Median_sal", `1st_sal`, `3rd_sal`, "Env_sal"
  )) |>
  mutate(
    var = ifelse(grepl("oxy", name), "Oxygen", NA),
    var = ifelse(grepl("sal", name), "Salinity", var),
    var = ifelse(grepl("temp", name), "Temperature", var),
    type = ifelse(grepl("Env", name), "Environment", "Biomass-weighted")
  ) |>
  filter(name %in% c("Median_oxy", "Median_temp", "Median_sal", "Env_oxy", "Env_temp", "Env_sal"))

t_env <- t |>
  filter(type == "Environment" & group2 == "Cod Adult") |>
  mutate(group2 = "Environment")

# Plot the slopes over time for group and quarter, compare to the environment slope.
s_slopes <- t |>
  filter(!type == "Environment") |>
  mutate(id = paste(group2, var, quarter, sep = "_")) |>
  ungroup() |>
  dplyr::select(year, id, value)

t_slopes <- t_env |>
  mutate(id = paste(group2, var, quarter, sep = "_")) |>
  dplyr::select(year, id, value)

slopes <- bind_rows(s_slopes, t_slopes)

# Calculate slopes over time
slopes2 <- slopes %>%
  split(.$id) |>
  purrr::map(~ lm(value ~ year, data = .x)) |>
  purrr::map_df(broom::tidy, .id = "Year") |>
  filter(term == "year") |>
  mutate(
    upr = estimate + 1.96 * std.error,
    lwr = estimate - 1.96 * std.error,
    sig = ifelse(estimate > lwr & estimate < upr, "sig.", "not sig.")
  ) |>
  rename(
    id = Year,
    year_slope = estimate
  ) |>
  dplyr::select(id, year_slope, sig, lwr, upr) |>
  separate(id, sep = "_", into = c("group2", "variable", "quarter"), remove = FALSE) |>
  mutate(quarter2 = ifelse(quarter == 1, "Quarter 1", "Quarter 4"))

vlines <- slopes2 |>
  filter(group2 == "Environment")

order <- c(
  "Cod Adult", "Cod Juvenile", "Flounder Adult", "Flounder Juvenile",
  "Plaice Adult", "Plaice Juvenile"
)

slopes2$group2 <- factor(slopes2$group2, levels = order)

p1 <- slopes2 |>
  filter(!group2 == "Environment") |>
  ggplot(aes(year_slope * 10, group2, xmin = lwr * 10, xmax = upr * 10, color = group2)) +
  geom_rect(
    data = vlines, aes(ymin = -Inf, ymax = Inf, xmin = lwr * 10, xmax = upr * 10, fill = "95% C.I. Environmental slope"),
    inherit.aes = FALSE, alpha = 0.1
  ) +
  geom_vline(xintercept = 0, linetype = 3, color = "tomato2", alpha = 1) +
  geom_vline(data = vlines, aes(xintercept = year_slope * 10, linetype = "Environmental slope"), alpha = 0.7) +
  geom_point(size = 2, position = position_dodge(width = 0.4)) +
  geom_errorbar(position = position_dodge(width = 0.4), width = 0) +
  facet_grid2(quarter2 ~ variable, scales = "free_x") +
  scale_fill_manual(values = "grey10") +
  scale_linetype_manual(values = 2) +
  scale_color_brewer(palette = "Paired") +
  labs(color = "", x = "Change in variable per decade", linetype = "", fill = "", y = "") +
  theme(
    legend.position = "bottom",
    strip.text.x.top = element_text(margin = margin(b = 2))
  ) +
  guides(color = "none") +
  NULL

tag_facet(p1, fontface = 1, col = "gray30", size = 3)

ggsave(paste0(home, "/figures/decadal_slopes.pdf"), width = 18, height = 11, units = "cm")
```

## Effective area occupied

```{r}
#| warning: false
#| message: false

eao_index_df <- eaos |>
  mutate(est_sc = est / mean(est), .by = group2) |>
  left_join(
    index |>
      dplyr::select(year, est, group2) |>
      rename(index = est) |>
      mutate(index_sc = index / max(index), .by = group2), ,
    by = c("year", "group2")
  )

# Slope + 95% CI of the linear est_sc ~ index_sc fit, per panel, to annotate
# alongside the geom_smooth(method = "lm") line already drawn on the plot.
slope_df <- eao_index_df |>
  reframe(broom::tidy(lm(est_sc ~ index_sc), conf.int = TRUE), .by = c(species, life_stage)) |>
  filter(term == "index_sc") |>
  mutate(label = sprintf("Slope: %.2f [%.2f, %.2f]", estimate, conf.low, conf.high))

p1 <- eao_index_df |>
  ggplot(aes(index_sc, est_sc, size = index_sc, color = year)) +
  geom_point(alpha = 0.8) +
  scale_color_viridis(end = 0.92, option = "magma") +
  scale_size_continuous(range = c(0.6, 4)) +
  guides(color = "none", size = "none") +
  geom_smooth(aes(index_sc, est_sc),
    se = FALSE,
    method = "gam", formula = y ~ s(x, k = 4),
    inherit.aes = FALSE, color = "steelblue2"
  ) +
  # geom_smooth(aes(index_sc, est_sc), inherit.aes = FALSE,
  #             se = FALSE, color = "steelblue2",
  #             method = "lm") +
  # geom_text(
  #   data = slope_df, aes(x = -Inf, y = Inf, label = label),
  #   inherit.aes = FALSE, hjust = -0.05, vjust = 3, size = 2.6, color = "grey20"
  # ) +
  labs(x = "Scaled biomass index", y = "Effective Area Occupied (relative to mean)") +
  facet_grid(life_stage ~ species)

tag_facet(p1, fontface = 1, col = "gray30", size = 3, x = Inf, hjust = 1.5)

ggsave(paste0(home, "/figures/eao_index.pdf"), width = 18, height = 11, units = "cm")
```

## Plot effects of area occupied by scenario (environment vs abundance)

```{r}
# Threshold is 5% of the median scenario's own max density, computed per group2
thresh_df <- scenario_pred |>
  dplyr::filter(scenario == "joint", level == "median") |>
  dplyr::summarise(thresh = 0.05 * max(est), .by = c(group2, species, life_stage))

scenario_labels <- c(
  environment = 'atop("Cool+High O"[2] * " " %->% " Warm+Low O"[2], "Abundance Held At Median")',
  index = 'atop("Low" * " " %->% " High Abundance", "Environment Held At Median")'
)

# Which level is the "from" (before) and "to" (after) endpoint, per scenario
transition_spec <- tibble::tribble(
  ~scenario, ~from, ~to,
  "environment", "high", "low",
  "index", "low", "high"
)

wide_df <- scenario_pred |>
  filter(level != "median", scenario %in% names(scenario_labels)) |>
  left_join(thresh_df, by = c("group2", "species", "life_stage")) |>
  mutate(occ = est >= thresh) |>
  dplyr::select(scenario, level, group2, species, life_stage, X, Y, occ) |>
  pivot_wider(names_from = level, values_from = occ, names_prefix = "occ_") |>
  left_join(transition_spec, by = "scenario") |>
  mutate(
    occ_from = if_else(from == "low", occ_low, occ_high),
    occ_to = if_else(to == "low", occ_low, occ_high),
    delta = case_when(
      occ_from & occ_to ~ "Core",
      occ_from & !occ_to ~ "Loss",
      !occ_from & occ_to ~ "Gain",
      TRUE ~ NA_character_
    )
  )

# Net change in occupied area from the transition described in the facet
# title -- the single number that answers "does occupied area go up or down"
delta_counts <- wide_df |>
  dplyr::summarise(
    n_core = sum(delta == "Core", na.rm = TRUE),
    n_appeared = sum(delta == "Gain", na.rm = TRUE),
    n_disappeared = sum(delta == "Loss", na.rm = TRUE),
    .by = c(group2, species, life_stage, scenario)
  ) |>
  dplyr::mutate(
    # Loss/gain expressed as % of the "from" area (n_core + n_disappeared),
    from_area = n_core + n_disappeared,
    loss_pct = 100 * n_disappeared / from_area,
    gain_pct = 100 * n_appeared / from_area,
    pct_change = gain_pct - loss_pct,
    # Turnover: total cells that changed (gained + lost) relative to the
    # stable core.Unlike pct_change, this doesn't let gains and losses
    # cancel out, so a driver that reshuffles a lot of area with
    # roughly-balanced gains/losses (small pct_change) still shows up here as
    # high turnover
    turnover_pct = 100 * (n_appeared + n_disappeared) / n_core
  )

ann_df <- dplyr::bind_rows(
  delta_counts |> dplyr::transmute(group2, species, life_stage, scenario,
    type = "Loss",
    label = paste0("Loss: -", round(loss_pct, 1), "%"), line = 1
  ),
  delta_counts |> dplyr::transmute(group2, species, life_stage, scenario,
    type = "Gain",
    label = paste0("Gain: +", round(gain_pct, 1), "%"), line = 2
  ),
  delta_counts |> dplyr::transmute(group2, species, life_stage, scenario,
    type = "net",
    label = paste0("Net ", ifelse(pct_change >= 0, "+", ""), round(pct_change), "%"),
    line = 3
  ),
  delta_counts |> dplyr::transmute(group2, species, life_stage, scenario,
    type = "turnover",
    label = paste0("Turnover: ", round(turnover_pct), "%"), line = 4
  )
) |>
  # Positive vjust: with y = Inf (top-left placement below), increasing vjust
  # pushes each line further down from the top edge. Spacing between lines is
  # the gap between these four values.
  dplyr::mutate(vjust = c(1.2, 2.3, 3.4, 4.5)[line])

pal <- RColorBrewer::brewer.pal(n = 9, name = "Spectral")

# delta_cols <- c(Core = "grey90", Loss = "tomato2", Gain = "steelblue")
delta_cols <- c(Core = alpha(pal[5], 0.5), Loss = pal[1], Gain = pal[9])
delta_legend_labels <- c(Core = "Occupied At Both", Loss = "Loss", Gain = "Gain")
text_cols <- c(Loss = "tomato2", Gain = "steelblue", net = "black", turnover = "black")

# One plot per life stage
plot_scenario_stage <- function(stage) {
  plot_map_fc +
    geom_tile(
      data = wide_df |>
        dplyr::filter(!is.na(delta), life_stage == stage) |>
        dplyr::mutate(X = X * 1000, Y = Y * 1000),
      aes(x = X, y = Y, fill = delta), width = 3000, height = 3000, inherit.aes = FALSE
    ) +
    facet_grid(species ~ scenario,
      labeller = labeller(scenario = as_labeller(scenario_labels, default = label_parsed))
    ) +
    scale_fill_manual(values = delta_cols, labels = delta_legend_labels, na.translate = FALSE) +
    scale_color_manual(values = text_cols) +
    guides(
      fill = guide_legend(override.aes = list(alpha = 1, colour = c("grey60", NA, NA), linewidth = 0.4)),
      color = "none"
    ) +
    geom_sf(size = 0.4, color = "grey80") +
    geom_label(
      data = ann_df |> dplyr::filter(life_stage == stage), aes(label = label, vjust = vjust),
      x = -Inf, y = Inf, hjust = -0.02, size = 3,
      linewidth = 0.2, label.padding = unit(0.1, "lines"),
      show.legend = FALSE, border.colour = "white",
    ) +
    labs(fill = NULL)
}

p1 <- plot_scenario_stage("Juvenile")
tag_facet(p1, fontface = 1, col = "gray20", size = 3, x = Inf, hjust = 1.2)
ggsave(paste0(home, "/figures/counterfact_juvenile.pdf"), width = 14, height = 18, units = "cm")

plot_scenario_stage("Adult")
tag_facet(p1, fontface = 1, col = "gray20", size = 3, x = Inf, hjust = 1.2)
ggsave(paste0(home, "/figures/counterfact_adult.pdf"), width = 14, height = 18, units = "cm")

# (p1 + p2) +
#   plot_annotation(tag_levels = "A") +
#   plot_layout(axes = "collect", guides = "collect") &
#   theme(legend.position = "bottom")
# ggsave(paste0(home, "/figures/test_adult.pdf"), width = 25, height = 18, units = "cm")
```

## Plot simulated occupancy transitions

```{r}
#| warning: false
#| message: false

order <- c(
  "Cod Adult", "Cod Juvenile", "Flounder Adult", "Flounder Juvenile",
  "Plaice Adult", "Plaice Juvenile"
)

occupancy_sims <- read_csv(paste0(home, "/output/occupancy_sims_02_sdms.csv")) |>
  separate(group, c("species", "life_stage"), sep = "_", remove = FALSE) |>
  mutate(
    species = str_to_title(species),
    life_stage = str_to_title(life_stage),
    group2 = factor(paste(species, life_stage), levels = order)
  ) |>
  pivot_longer(c(gain, loss, net, turnover), names_to = "metric", values_to = "value") |>
  mutate(metric = str_to_title(metric)) |>
  filter(metric %in% c("Net", "Turnover")) |>
  filter(value < 375)

# Boxplot uses the full 500-draw sample; the overlaid jitter is a 50-point
# subsample per group/scenario/metric
occupancy_sims_sub <- occupancy_sims |>
  filter(metric %in% c("Net", "Turnover")) |>
  slice_sample(n = 50, by = c(group2, scenario, metric)) |>
  filter(value < 375)

net_pos_prop <- occupancy_sims |>
  filter(metric == "Net") |>
  summarise(prop_pos = mean(value > 0), x_max = max(value), .by = c(group2, scenario)) |>
  mutate(
    prop_neg = 1 - prop_pos,
    label = paste0(round(100 * prop_neg), "/", round(100 * prop_pos)),
    metric = "Net"
  )

p1 <- occupancy_sims |>
  ggplot(aes(value, group2, color = scenario)) +
  geom_boxplot(
    outlier.color = NA, size = 0.5, alpha = 0.65, width = 0.4,
    position = position_dodge(width = 0.5)
  ) +
  geom_jitter(
    data = occupancy_sims_sub,
    position = position_jitterdodge(dodge.width = 0.5, jitter.width = 0.4, jitter.height = 0),
    stroke = 1, alpha = 0.2, size = 0.1
  ) +
  geom_vline(
    data = tibble(metric = "Net"), aes(xintercept = 0),
    linetype = 2, alpha = 0.5, linewidth = 0.4
  ) +
  geom_text(
    data = net_pos_prop,
    aes(x = 230, y = group2, label = label, color = scenario),
    position = position_dodge(width = 0.75), hjust = -0.1, size = 2.8, show.legend = FALSE
  ) +
  ggh4x::facet_grid2(scenario ~ metric,
    scales = "free_x",
    labeller = labeller(scenario = as_labeller(scenario_labels, default = label_parsed))
  ) +
  scale_color_brewer(palette = "Dark2") +
  guides(color = "none") +
  theme(
    strip.text.y.right = element_text(size = 9.4),
    strip.text.x.top = element_text(size = 9.4)
  ) +
  labs(x = "Change in occupied area (%)", y = "")

tag_facet(p1, fontface = 1, col = "gray30", size = 3, x = Inf, hjust = 1.5)

ggsave(paste0(home, "/figures/occupancy_sims.pdf"), width = 18, height = 12, units = "cm")
```

## Plot model coefficients

```{r}
#| warning: false
#| message: false

term_labels <- c(
  sal_sc = "Salinity",
  depth_sc = "Depth",
  depth_sq = "Depth²",
  temp_sc = "Temperature",
  temp_sq = "Temperature²",
  "oxy_sc-slope" = "Oxygen (slope)",
  "oxy_sc-breakpt" = "Oxygen (breakpoint)",
  sigma_O = "Spatial SD (σO)",
  sigma_E = "Spatiotemporal SD (σE)",
  sigma_Z = "SVC SD (σZ)"
)

order <- c(
  "Cod Adult", "Cod Juvenile", "Flounder Adult", "Flounder Juvenile",
  "Plaice Adult", "Plaice Juvenile"
)

term_exprs <- c(
  Salinity = "Salinity",
  Depth = "Depth",
  "Depth²" = "Depth^2",
  Temperature = "Temperature",
  "Temperature²" = "Temperature^2",
  "Oxygen (slope)" = '"Oxygen (slope)"',
  "Oxygen (breakpoint)" = '"Oxygen (breakpoint)"',
  "Spatial SD (σO)" = '"Spatial SD ("*sigma[O]*")"',
  "Spatiotemporal SD (σE)" = '"Spatiotemporal SD ("*sigma[E]*")"',
  "SVC SD (σZ)" = '"SVC SD ("*sigma[Z]*")"'
)

coefs <- read_csv(paste0(home, "/output/coefs_02_sdms.csv")) |>
  filter(term %in% names(term_labels)) |>
  separate(group, c("species", "life_stage"), sep = "_", remove = FALSE) |>
  mutate(
    species = str_to_title(species),
    life_stage = str_to_title(life_stage),
    group2 = factor(paste(species, life_stage), levels = order),
    term_label = factor(term_labels[term], levels = term_labels)
  )

coefs |>
  ggplot(aes(estimate, term_label, xmin = conf.low, xmax = conf.high, color = group2)) +
  geom_stripped_rows(aes(y = term_label), inherit.aes = FALSE) +
  geom_vline(xintercept = 0, linetype = 3, color = "grey30", alpha = 1) +
  geom_point(position = position_dodge(width = 0.7)) +
  geom_errorbar(width = 0, position = position_dodge(width = 0.7), alpha = 0.4) +
  scale_color_brewer(palette = "Paired") +
  scale_y_discrete(labels = function(x) parse(text = term_exprs[x])) +
  labs(color = "", y = "", x = "Estimate") +
  theme(legend.position = "bottom")

ggsave(paste0(home, "/figures/coef_caterpillar.pdf"), width = 18, height = 11, units = "cm")
```
