---
title: "Plot conditional effects and grid predictions from 02"
author: "Max Lindmark"
date: today
date-format: iso
toc: true
format: 
  html:
    page-layout: full
    embed-resources: true
knitr: 
  opts_chunk:
    fig.align: center
    out-width: 100%
editor: source
---

## Intro

Plot conditional effects from predictions in script "02-fit-sdms-conditional.qmd"

## Load packages & source functions

```{r load libraries}
#| message: false
#| warning: false

# Load libraries, install if needed
pkgs <- c("tidyverse", "readxl", "tidylog", "RCurl", "devtools", "patchwork",
          "viridis", "RColorBrewer", "here", "ggnewscale", "ggh4x") 

if(length(setdiff(pkgs,rownames(installed.packages()))) > 0){

    install.packages(setdiff(pkgs, rownames(installed.packages())), dependencies = T)
  
  }

invisible(lapply(pkgs, library, character.only = T))

# Source code for map plots
# You need: # devtools::install_github("seananderson/ggsidekick") # not on CRAN; library(ggsidekick)
devtools::source_url("https://raw.githubusercontent.com/maxlindmark/spatial-metabolic-index/main/R/functions/map-plot.R")
options(ggplot2.continuous.colour = "viridis")

pal_temp <- brewer.pal(n = 10, name = "Paired")[c(2, 6)]
pal_oxy <- brewer.pal(n = 10, name = "Paired")[c(8, 4)]

# Set path
home <- here::here()
```

## Read predictions

```{r}
preds <- read_csv(paste0(home, "/output/conditional_02_sdms.csv")) %>% 
  filter(oxy > 3 & oxy < 9) %>%
  filter(temp > 2 & temp < 14) %>% 
  separate(group, c("species", "life_stage"), sep = "_", remove = FALSE) %>% 
  mutate(species = str_to_title(species),
         life_stage = str_to_title(life_stage)) %>% 
  group_by(group) %>% 
  mutate(est_sc = exp(est)/max(exp(est))) # for plotting all together in heatmap

grid_preds <- read_csv(paste0(home, "/output/pred_grids_02_sdms.csv")) %>% 
  separate(group, c("species", "life_stage"), sep = "_", remove = FALSE) %>% 
  mutate(species = str_to_title(species),
         life_stage = str_to_title(life_stage))

cogs <- read_csv(paste0(home, "/output/cogs_02_sdms.csv")) %>% 
  separate(group, c("species", "life_stage"), sep = "_", remove = FALSE) %>% 
  mutate(species = str_to_title(species),
         life_stage = str_to_title(life_stage))
```

## Calculate weighted quantiles

```{r}
wq <- grid_preds %>% 
  group_by(group, quarter) %>%
  summarise("Median_oxy" = hutils::weighted_quantile(v = oxy, w = exp(est), p = 0.5),
            "10_oxy" = hutils::weighted_quantile(v = oxy, w = exp(est), p = 0.1),
            "25_oxy" = hutils::weighted_quantile(v = oxy, w = exp(est), p = 0.25),
            "75_oxy" = hutils::weighted_quantile(v = oxy, w = exp(est), p = 0.75),
            "90_oxy" = hutils::weighted_quantile(v = oxy, w = exp(est), p = 0.9),
            "Env_oxy" = median(oxy),
            
            "Median_temp" = hutils::weighted_quantile(v = temp, w = exp(est), p = 0.5),
            "10_temp" = hutils::weighted_quantile(v = temp, w = exp(est), p = 0.1),
            "25_temp" = hutils::weighted_quantile(v = temp, w = exp(est), p = 0.25),
            "75_temp" = hutils::weighted_quantile(v = temp, w = exp(est), p = 0.75),
            "90_temp" = hutils::weighted_quantile(v = temp, w = exp(est), p = 0.9),
            "Env_temp" = median(temp)) %>% 
  ungroup() %>% 
  separate(group, c("species", "life_stage"), sep = "_", remove = FALSE) %>% 
  mutate(species = str_to_title(species),
         life_stage = str_to_title(life_stage)) 

wq_annual <- grid_preds %>% 
  group_by(group, year, quarter) %>%
  # Could use reframe here... 
  summarise("Median_oxy" = hutils::weighted_quantile(v = oxy, w = exp(est), p = 0.5),
            "1st_oxy" = hutils::weighted_quantile(v = oxy, w = exp(est), p = 0.25),
            "3rd_oxy" = hutils::weighted_quantile(v = oxy, w = exp(est), p = 0.75),
            "Env_oxy" = median(oxy),
            
            "Median_temp" = hutils::weighted_quantile(v = temp, w = exp(est), p = 0.5),
            "1st_temp" = hutils::weighted_quantile(v = temp, w = exp(est), p = 0.25),
            "3rd_temp" = hutils::weighted_quantile(v = temp, w = exp(est), p = 0.75),
            "Env_temp" = median(temp)) %>% 
  ungroup() %>% 
  separate(group, c("species", "life_stage"), sep = "_", remove = FALSE) %>% 
  mutate(species = str_to_title(species),
         life_stage = str_to_title(life_stage)) 
```

## Plot conditional effects

```{r}
# Oxygen
preds_oxy <- preds %>% 
  filter(temp %in% sort(unique(preds$temp))[0.5*length(unique(preds$temp))])

wi_range_oxy <- preds_oxy %>% 
  left_join(wq %>%
              filter(quarter == 4) %>%
              dplyr::select(group, `10_oxy`, `25_oxy`, `75_oxy`, `90_oxy`), by = "group")

# TODO: make predictions for the exact value??
wi_median_oxy <- preds_oxy %>% 
  left_join(wq %>%
              filter(quarter == 4) %>%
              dplyr::select(group, Median_oxy), by = "group") %>% 
  filter(oxy > 0.985*Median_oxy & oxy < 1.015*Median_oxy) %>% 
  as.data.frame()

col_oxy <- "grey30"#viridis(n = 12, option = "mako")[7]
col_temp <- "grey30"#viridis(n = 12, option = "plasma")[8]

p1 <- ggplot() + 
  ggh4x::facet_grid2(life_stage ~ species, scales = "free", independent = "y") +
  # NOTE: 80% Confidence interval
  
  # 25th and 75th percentile
  geom_ribbon(data = wi_range_oxy %>% filter(oxy > `25_oxy` & oxy < `75_oxy`), fill = col_oxy,
              aes(oxy, exp(est), ymin = exp(est - est_se*1.282), ymax = exp(est + est_se*1.282)),
              color = NA, alpha = 0.2) +
  geom_line(data = wi_range_oxy %>% filter(oxy > `25_oxy` & oxy < `75_oxy`), aes(oxy, exp(est)),
            color = col_oxy, alpha = 0.5, linewidth = 1.5) + 
  
  # 10th and 90th percentile
  geom_ribbon(data = wi_range_oxy %>% filter(oxy > `10_oxy` & oxy < `90_oxy`), fill = col_oxy, 
              aes(oxy, exp(est), ymin = exp(est - est_se*1.282), ymax = exp(est + est_se*1.282)),
              color = NA, alpha = 0.2) +
  geom_line(data = wi_range_oxy %>% filter(oxy > `10_oxy` & oxy < `90_oxy`), aes(oxy, exp(est)),
            color = col_oxy, alpha = 0.5, linewidth = 1) + 
  
  # NOTE: 85% Confidence interval
  geom_ribbon(data = preds_oxy, fill = col_oxy,
              aes(oxy, exp(est), ymin = exp(est - est_se*1.282), ymax = exp(est + est_se*1.282)),
              color = NA, alpha = 0.2) +
  geom_line(data = preds_oxy, aes(oxy, exp(est)), alpha = 0.5, linewidth = 0.5, color = col_oxy) + 
  
  # Median!
  geom_point(data = wi_median_oxy, aes(Median_oxy, exp(est)),
             color = col_oxy, alpha = 0.5, size = 2.5) + 
  
  labs(x = "Oxygen (ml/L)", y = "Biomass density (kg/km<sup>2</sup>)") +  
  theme(#legend.position = "bottom",
        #legend.spacing.y = unit(0.1, 'cm'),
        #legend.key = element_rect(fill = "grey95", inherit.blank = TRUE),
        axis.title.y = ggtext::element_markdown(),
        plot.margin = unit(c(0,0,0,0), "cm")) +
  NULL

# Temperature
preds_temp <- preds %>% 
  filter(oxy %in% sort(unique(preds$oxy))[0.5*length(unique(preds$oxy))])

wi_range_temp <- preds_temp %>% 
  left_join(wq %>%
              filter(quarter == 4) %>%
              dplyr::select(group, `10_temp`, `25_temp`, `75_temp`, `90_temp`), by = "group")

# TODO: make predictions for the exact value??
wi_median_temp <- preds_temp %>% 
  left_join(wq %>%
              filter(quarter == 4) %>%
              dplyr::select(group, Median_temp), by = "group") %>% 
  filter(temp > 0.985*Median_temp & temp < 1.015*Median_temp) %>% 
  as.data.frame()

wi_median_temp

p2 <- ggplot() + 
  ggh4x::facet_grid2(life_stage ~ species, scales = "free", independent = "y") +
  # NOTE: 80% Confidence interval
  
  # 25th and 75th percentile
  geom_ribbon(data = wi_range_temp %>% filter(temp > `25_temp` & temp < `75_temp`), fill = col_temp,
              aes(temp, exp(est), ymin = exp(est - est_se*1.282), ymax = exp(est + est_se*1.282)), 
              color = NA, alpha = 0.2) +
  geom_line(data = wi_range_temp %>% filter(temp > `25_temp` & temp < `75_temp`), aes(temp, exp(est)),
            color = col_temp, alpha = 0.5, linewidth = 1.5) + 
   
  # 10th and 90th percentile
  geom_ribbon(data = wi_range_temp %>% filter(temp > `10_temp` & temp < `90_temp`), fill = col_temp, 
              aes(temp, exp(est), ymin = exp(est - est_se*1.282), ymax = exp(est + est_se*1.282)),
              color = NA, alpha = 0.2) +
  geom_line(data = wi_range_temp %>% filter(temp > `10_temp` & temp < `90_temp`), aes(temp, exp(est)),
            color = col_temp, alpha = 0.5, linewidth = 1) + 
  
  # NOTE: 85% Confidence interval
  geom_ribbon(data = preds_temp, fill = col_temp,
              aes(temp, exp(est), ymin = exp(est - est_se*1.282), ymax = exp(est + est_se*1.282)),
              color = NA, alpha = 0.2) +
  geom_line(data = preds_temp, aes(temp, exp(est)), alpha = 0.5, linewidth = 0.5, color = col_temp) + 
  
  # Median!
  geom_point(data = wi_median_temp, aes(Median_temp, exp(est)),
             color = col_temp, alpha = 0.5, size = 2.5) + 
  
  labs(x = "Temperature (°C)", y = "Biomass density (kg/km<sup>2</sup>)") +  
  theme(#legend.position = "bottom",
        #legend.spacing.y = unit(0.1, 'cm'),
        #legend.key = element_rect(fill = "grey95", inherit.blank = TRUE),
        axis.title.y = ggtext::element_markdown(),
        plot.margin = unit(c(0,0,0,0), "cm")) +
  NULL

(p1 / p2) + plot_annotation(tag_levels = "A")

ggsave(paste0(home, "/figures/oxy_temp_conditional.pdf"), width = 17, height = 17, units = "cm")
```

## Plot weighted quantiles over time

```{r}
t <- wq_annual %>% 
  pivot_longer(c("Median_oxy", `1st_oxy`, `3rd_oxy`, "Env_oxy", "Median_temp", `1st_temp`, `3rd_temp`, "Env_temp")) %>% 
  mutate(var = ifelse(grepl("oxy", name), "Oxygen", "Temperature"),
         type = ifelse(grepl("Env", name), "Environment", "Biomass-weighted")) %>% 
  separate(group, c("species", "life_stage"), sep = "_", remove = FALSE) %>% 
  mutate(species = str_to_title(species),
         life_stage = str_to_title(life_stage),
         group2 = paste(species, life_stage, sep = " ")) %>% 
  filter(name %in% c("Median_oxy", "Median_temp", "Env_oxy", "Env_temp"))
  
t_env <- t %>%
  filter(type == "Environment" & group == "cod_adult") %>% 
  mutate(group2 = "Environment")

t2 <- bind_rows(t_env, t %>% filter(!type == "Environment"))

# Reorder to have Environment at the end
t2$group2 <- factor(t2$group2, levels = rev(unique(t2$group2)))

pal <- brewer.pal(name = "Dark2", n = 8)[3:8]

t2 %>% 
  mutate(quarter2 = ifelse(quarter == 1, "Quarter 1", "Quarter4")) %>% 
  ggplot(aes(year, value, color = group2, linetype = group2)) +
  geom_line(alpha = 0.8) +
  #facet_grid(quarter ~ var) +
  ggh4x::facet_grid2(var ~ quarter2, scales = "free") +
  labs(y = "Environmental variable", x = "Year", linetype = "", color = "") +
  guides(color = guide_legend(nrow = 3,override.aes = list(linetype = c(rep(1, 6), 2), 
                                                           size = 0.5)),
         linetype = "none") +
  scale_x_continuous(breaks = c(seq(1993, 2021, by = 5))) +
  scale_linetype_manual(values = c(rep(1, 6), 2)) +
  scale_color_manual(values = c(rev(pal), "grey40")) +
  theme(legend.position = c(0.24, 0.36),
        legend.spacing.y = unit(-2, 'cm'),
        legend.spacing.x = unit(0, 'cm'),
        legend.background = element_rect(fill = NA))

ggsave(paste0(home, "/figures/wm_temp_oxy.pdf"), width = 17, height = 14, units = "cm")
```

## Plot centrs of gravity

```{r}
plot_map +
  geom_point(data = cogs, aes(est_x, est_y, color = year), size = 0.5) +
  geom_errorbar(data = cogs, aes(y = est_y, xmin = lwr_x, xmax = upr_x, color = year),
                width = 0, alpha = 0.3, linewidth = 0.3) +
  geom_errorbar(data = cogs, aes(x = est_x, ymin = lwr_y, ymax = upr_y, color = year),
                width = 0, alpha = 0.3, linewidth = 0.3) +
  geom_path(data = cogs, aes(est_x, est_y, color = year), linewidth = 0.3) + 
  facet_grid(life_stage ~ species) + 
  labs(color = "Year") +
  theme(legend.position = c(0.05, 0.85),
        legend.key.height = unit(0.3, "cm"),
        legend.key.width = unit(0.1, "cm"))

ggsave(paste0(home, "/figures/cogs_map.pdf"), width = 17, height = 11, units = "cm")

pal <- brewer.pal(name = "Dark2", n = 8)[3:8]

cogs <- cogs %>% 
  mutate(group2 = paste(species, life_stage, sep = " "))

p1 <- ggplot(cogs, aes(year, est_x, color = group2, fill = group2)) + 
  geom_ribbon(aes(ymin = lwr_x, ymax = upr_x), color = NA, alpha = 0.3) + 
  geom_line() + 
  scale_color_manual(values = c(pal)) + 
  scale_fill_manual(values = c(pal)) + 
  guides(color = "none", fill = "none") +
  labs(x = "", y = "UTM X center of gravity", color = "", fill = "")

p2 <- ggplot(cogs, aes(year, est_y, color = group2, fill = group2)) + 
  geom_ribbon(aes(ymin = lwr_y, ymax = upr_y), color = NA, alpha = 0.3) + 
  geom_line() + 
  scale_color_manual(values = c(pal)) + 
  scale_fill_manual(values = c(pal)) + 
  guides(color = guide_legend(nrow = 2)) +
  labs(x = "", y = "UTM Y center of gravity", color = "", fill = "") +
  theme(legend.position = "bottom",
        legend.spacing.y = unit(-2, 'cm'),
        legend.spacing.x = unit(0, 'cm'),
        legend.background = element_rect(fill = NA))

p1 / p2

ggsave(paste0(home, "/figures/supp/cogs.pdf"), width = 17, height = 20, units = "cm")
```

## Correlation and slope plot

```{r}
#  Correlation plot
dcor <- t %>% filter(!type == "Environment") %>%
  mutate(id = paste(name, year, quarter, sep = "_")) %>%
  dplyr::select(group2, year, var, value, quarter, id)

t_env2 <- t_env %>%
  mutate(name = ifelse(name == "Env_oxy", "Median_oxy", "Median_temp")) %>%
 mutate(id = paste(name, year, quarter, sep = "_")) %>%
  rename(env_value = value) %>% 
  dplyr::select(id, env_value)

dcor2 <- dcor %>% left_join(t_env2, by = "id")
 
cors <- plyr::ddply(dcor2, c("group2", "var", "quarter"),
                    summarise, cor = round(cor(env_value, value), 2)) %>% 
  arrange(var)

hist(cors$cor)

# Plot correlations between environment and weighted
cor_pal <- brewer.pal(n = 8, name = "Dark2")[6:7]
  
dcor2 %>% 
  mutate(quarter2 = ifelse(quarter == 1, "Quarter 1", "Quarter 4")) %>% 
  ggplot(aes(env_value, value, color = quarter2)) +
  ggh4x::facet_grid2(var ~ group2, scales = "free") +
  geom_abline(color = "grey30", linetype = 2, linewidth = 0.5) +
  geom_point(alpha = 0.7, size = 0.5) +
  geom_text(data = cors %>% filter(quarter == 1),
            aes(label = paste("r=", cor, sep = "")), x = -Inf, y = Inf, hjust = -0.1, vjust = 1.5,
            inherit.aes = FALSE, fontface = "italic", color = cor_pal[2], size = 2.5) +
  geom_text(data = cors %>% filter(quarter == 4),
            aes(label = paste("r=", cor, sep = "")), x = -Inf, y = Inf, hjust = -0.1, vjust = 2.7,
            inherit.aes = FALSE, fontface = "italic", color = cor_pal[1], size = 2.5) +
  labs(y = "Biomass-weighted", x = "Environment", color = "") +
  stat_smooth(method = "lm", se = FALSE, linewidth = 0.5) +
  scale_color_manual(values = cor_pal) +
  scale_x_continuous(breaks = c(4:8)) +
  theme_sleek(base_size = 10) + 
  theme(aspect.ratio = 1,
        legend.key.size = unit(0.2, 'cm'),
        legend.text = element_text(size = 6),
        legend.position = c(0.11, 0),
        legend.spacing.y = unit(-1, 'cm'))

ggsave(paste0(home, "/figures/supp/wm_cors.pdf"), width = 17, height = 7, units = "cm")

# Plot the slopes over time for group and quarter, compare to the environment slope. Boxplots and lines??
s_slopes <- t %>% filter(!type == "Environment") %>%
  mutate(id = paste(group, var, quarter, sep = "_")) %>% 
  ungroup() %>% 
  dplyr::select(year, id, value)

t_slopes <- t_env %>%
  mutate(id = paste("Environment", group2, var, quarter, sep = "_")) %>% 
  dplyr::select(year, id, value)

slopes <- bind_rows(s_slopes, t_slopes)

# Calculate slopes over time
slopes2 <- slopes %>%
  split(.$id) %>%
  purrr::map(~lm(value ~ year, data = .x)) %>%
  purrr::map_df(broom::tidy, .id = 'Year') %>%
  filter(term == "year") %>% 
  mutate(upr = estimate + 1.96*std.error,
         lwr = estimate - 1.96*std.error,
         sig = ifelse(estimate > lwr & estimate < upr, "sig.", "not sig.")) %>% 
  rename(id = Year, 
         year_slope = estimate) %>% 
  dplyr::select(id, year_slope, sig, lwr, upr) %>% 
  separate(id, into = c("species", "life_stage", "variable", "quarter"), remove = FALSE) %>% 
  #mutate(group3 = paste(str_to_title(species), str_to_title(life_stage), paste0("Q", quarter)))
  mutate(group3 = paste(str_to_title(species), str_to_title(life_stage)),
         x = "x") %>% 
  mutate(quarter2 = ifelse(quarter == 1, "Quarter 1", "Quarter 4"))
  
hlines <- slopes2 %>% 
  filter(species == "Environment")

pal <- brewer.pal(name = "Dark2", n = 8)[3:8]

slopes2 %>% 
  filter(!species == "Environment") %>% 
  ggplot(aes(x, year_slope, ymin = lwr, ymax = upr, color = group3)) + 
  geom_rect(data = hlines, aes(xmin = -Inf, xmax = Inf, ymin = lwr, ymax = upr, fill = "Env. slope 95% CI"),
            inherit.aes = FALSE, alpha = 0.1) +
  geom_hline(yintercept = 0, linetype = 3, color = "tomato2", alpha = 1) +
  geom_hline(data = hlines, aes(yintercept = year_slope, linetype = "Env. slope"), alpha = 0.7) +
  geom_point(size = 2.5, position = position_dodge(width = 0.4)) + 
  geom_errorbar(position = position_dodge(width = 0.4), width = 0) +
  ggh4x::facet_grid2(variable ~ quarter2, scales = "free") +
  scale_fill_manual(values = "grey10") +
  scale_linetype_manual(values = 2) +
  scale_color_manual(values = pal) +
  labs(color = "", y = "Year-slope", linetype = "", fill = "") +
  theme(axis.text.x = element_blank(),
        axis.ticks.x = element_blank(),
        axis.title.x = element_blank(), 
        legend.position = "bottom") +
  NULL

ggsave(paste0(home, "/figures/year_slopes.pdf"), width = 17, height = 17, units = "cm")
```

## Plot spatially-varying quarter effects, spatial predictions, and ST random fields

```{r}
grid_preds <- grid_preds %>% 
  mutate(group2 = paste(species, life_stage, sep = " ")) 
  
names(grid_preds)

# Plot spatially-varying effect
sv <- grid_preds %>% 
  #filter(group == pull(grid_preds, group)[1]) %>% 
  filter(year == 1999) %>% 
  pivot_longer(c("zeta_s_quarter_f1", "zeta_s_quarter_f4"),
               names_to = "field", values_to = "zeta") %>% 
  mutate(Quarter = ifelse(field == "zeta_s_quarter_f1", "Quarter 1", "Quarter 4"))

plot_map_fc +
  geom_raster(data = sv, aes(X*1000, Y*1000, fill = zeta)) + 
  facet_grid(Quarter ~ group2) +
  scale_fill_gradient2(name = "Zeta") + 
  theme_sleek(base_size = 8) +
  theme(axis.text.x = element_text(angle = 90),
        legend.position = "bottom",
        legend.key.height = unit(0.2, "cm"))

ggsave(paste0(home, "/figures/supp/sv_effects.pdf"), width = 17, height = 9, units = "cm")


for(i in unique(grid_preds$group2)){
  
  dd <- grid_preds %>% filter(group2 == i)
  j <- pull(dd, group)[1]
  
  # Quarter 1
  print(
    plot_map_fc +
      geom_raster(data = filter(dd, quarter == 1), aes(X*1000, Y*1000, fill = exp(est))) +
      facet_wrap(~year) + 
      scale_fill_viridis(trans = "sqrt", name = "Biomass density (kg/km)",
                         # trim extreme high values to make spatial variation more visible
                         na.value = "yellow", limits = c(0, exp(quantile(filter(dd, quarter == 4)$est, 0.999)))) +
      labs(caption = paste("maximum estimated biomass density =", round(max(exp(filter(dd, quarter == 4)$est)))),
           title = i, subtitle = "Quarter 1")
    )
    
    ggsave(paste0(home, paste0("/figures/supp/spatial_biomass", "_q1_", j, ".pdf")), width = 17, height = 17, units = "cm")
    
    
  # Quarter 4
  print(
    plot_map_fc +
      geom_raster(data = filter(dd, quarter == 4), aes(X*1000, Y*1000, fill = exp(est))) +
      facet_wrap(~year) + 
      scale_fill_viridis(trans = "sqrt", name = "Biomass density (kg/km)",
                         # trim extreme high values to make spatial variation more visible
                         na.value = "yellow", limits = c(0, exp(quantile(filter(dd, quarter == 4)$est, 0.999)))) +
      labs(caption = paste("maximum estimated biomass density =", round(max(exp(filter(dd, quarter == 4)$est)))),
           title = i, subtitle = "Quarter 4")
    )
    
    ggsave(paste0(home, paste0("/figures/supp/spatial_biomass", "_q4_", j, ".pdf")), width = 17, height = 17, units = "cm")
  
}
```
