---
title: "Fit random forest models to spatial biomass trends and velocities"
author: "Max Lindmark"
date: today
date-format: iso
toc: true
format: 
  html:
    page-layout: full
    embed-resources: true
knitr: 
  opts_chunk:
    fig.align: center
    out-width: 100%
editor: source
---

## Intro

Fit random forest models to biomass trends and velocities from sdms random fields. We'll use a `tidymodels` pipeline, consisting of `rsample` for sample splitting (e.g. train/test or cross-validation), `recipes` for pre-processing, `parsnip` for specifying the model and `yardstick` for evaluating the model.

## Load packages & source functions

```{r load libraries}
#| message: false
#| warning: false

library(tidyverse)
library(tidylog)
library(tidymodels)
library(RCurl)
library(devtools)
library(vip)
library(viridis)
library(RColorBrewer)
library(ranger)
library(patchwork)
library(DALEXtra)
library(modelr)

# Source code for map plots
devtools::source_url("https://raw.githubusercontent.com/maxlindmark/spatial-metabolic-index/main/R/functions/map-plot.R")

# Set path
home <- here::here()

# Set seed
set.seed(42)
```

## Load data

```{r}
#| message: false
#| warning: false

d_old <- read_csv(paste0(home, "/output/rf_dat_long.csv")) |> 
  filter(cumulative == "Y") |> 
  drop_na(temp_vel, oxy_vel) |> 
  dplyr::select(biom_trend, temp_vel, mean_temp, oxy_vel, mean_oxy, group, mean_biom)

d_short <- read_csv(paste0(home, "/output/rf_dat_short.csv")) |> 
  filter(cumulative == "Y") |> 
  drop_na(temp_vel, oxy_vel) |> 
  dplyr::select(biom_trend, temp_vel, mean_temp, oxy_vel, mean_oxy, group, mean_biom)
```

```{r}
#| include: false
# Some resources
# https://brunaw.com/slides/rladies-dublin/RF/intro-to-rf.html#27
# https://www.rebeccabarter.com/blog/2020-03-25_machine_learning#what-is-tidymodels
# https://juliasilge.com/blog/ikea-prices/
# https://juliasilge.com/blog/sf-trees-random-tuning/
# https://www.tidymodels.org/start/tuning/
# https://workflows.tidymodels.org/reference/predict-workflow.html
# https://christophm.github.io/interpretable-ml-book/feature-importance.html 
```

## Biomass trends with trends-covariates
First we split the data into training and testing (for evaluating the final model's performance after tuning hyper-parameters using cross validation)

```{r}
trend_old <- d_old

trend_split <- initial_split(trend_old, prop = 3/4)
trend_split

# Extract training and testing sets
trend_train <- training(trend_split)
trend_test <- testing(trend_split)

# We need to tune the model by comparing the predictive performance of hyper parameters using cross validation, so set up a cross-validation object.

# Create CV object from training data
trend_cv <- vfold_cv(trend_train)

# Now define the "recipe", basically the formula because we have already pre-processed the data

# Define the recipe, note the variables are already scaled
trend_recipe <- recipe(biom_trend ~ temp_vel + mean_temp + oxy_vel + mean_oxy + group + mean_biom,
                       data = trend_old)

trend_recipe

# Next step is to pre-specify the model using `parsnip` It's basically a "universal" way to specify a model by defining the model type, arguments (such as which hyper parameters to tune), engine. We tune mtry (number of predictors that are sampled at splits in a tree-based model node) and trees (how many models to average predictions over)

rf_trend_model <- 
  # specify that the model is a random forest
  rand_forest() |>
  # specify what hyperparameters to tune
  set_args(mtry = tune(),
           trees = tune()) |>
  # select the engine/package that underlies the model
  set_engine("ranger", importance = "permutation", keep.inbag = TRUE) |> # keep.inbag is to get CI's later!
  set_mode("regression")

rf_trend_model

# Put the model and recipes together into a workflow. You initiate a workflow using `workflow()`, and then you can add a recipe and add a model to it.

# Set the workflow
rf_trend_workflow <- workflow() |>
  add_recipe(trend_recipe) |>
  add_model(rf_trend_model)

rf_trend_workflow

# Tune hyper parameters by checking model performance using cross validation

# Default mtry is the (rounded down) square root of the number variables and tree is 500. Specify which values want to try in a grid
rf_trend_grid <- expand.grid(
  mtry = seq(2, 5, by = 1),
  trees = seq(200, 1000, by = 200)
  )

# Extract results of tuning (metric is rmse)
rf_trend_tune_results <- rf_trend_workflow |>
  tune_grid(resamples = trend_cv, # CV object
            grid = rf_trend_grid) # grid of values to try

# Print results
rf_trend_tune_results |>
  collect_metrics()

p1 <- autoplot(rf_trend_tune_results) +
  scale_color_brewer(palette = "Set1") + 
  theme(legend.position = "bottom")
# Seems mtry = 3 is the best (just) 

p1 

# Now update the workflow to get the best mtry and n trees from the tuning
param_trend_final <- rf_trend_tune_results |>
  select_best(metric = "rmse")

param_trend_final

# Add this parameter to the workflow using the finalize_workflow() function.
rf_trend_workflow <- rf_trend_workflow |>
  finalize_workflow(param_trend_final) # can do the select best in here as well

rf_trend_workflow

# Now we’ve defined our recipe, our model, and tuned the model’s parameters, we can fit the final model. Applying the `last_fit()` function to our workflow and our train/test split object will train the model specified by the workflow using the training data, and produce evaluations based on the test set.

rf_trend_fit_old <- rf_trend_workflow |>
  last_fit(trend_split)

# Evaluate the model performance on the testing data

rf_trend_fit_old |> collect_metrics()

# Now that we have a final model we want to train it on the full dataset, and then use it to predict the response for new data. For that we need to `fit()` again on the full dataset (the complete training + testing dataset).

final_trend_model <- fit(rf_trend_workflow, trend_old)

final_trend_model

# Save!
saveRDS(final_trend_model, paste0(home, "/output/rf_final_trend_model_old.rds"))
```

```{r}
trend_short <- d_short

trend_split <- initial_split(trend_short, prop = 3/4)
trend_split

# Extract training and testing sets
trend_train <- training(trend_split)
trend_test <- testing(trend_split)

# We need to tune the model by comparing the predictive performance of hyper parameters using cross validation, so set up a cross-validation object.

# Create CV object from training data
trend_cv <- vfold_cv(trend_train)

# Now define the "recipe", basically the formula because we have already pre-processed the data

# Define the recipe, note the variables are already scaled
trend_recipe <- recipe(biom_trend ~ temp_vel + mean_temp + oxy_vel + mean_oxy + group + mean_biom,
                       data = trend_short)

trend_recipe

# Next step is to pre-specify the model using `parsnip` It's basically a "universal" way to specify a model by defining the model type, arguments (such as which hyper parameters to tune), engine. We tune mtry (number of predictors that are sampled at splits in a tree-based model node) and trees (how many models to average predictions over)

rf_trend_model <- 
  # specify that the model is a random forest
  rand_forest() |>
  # specify what hyperparameters to tune
  set_args(mtry = tune(),
           trees = tune()) |>
  # select the engine/package that underlies the model
  set_engine("ranger", importance = "permutation", keep.inbag = TRUE) |> # keep.inbag is to get CI's later!
  set_mode("regression")

rf_trend_model

# Put the model and recipes together into a workflow. You initiate a workflow using `workflow()`, and then you can add a recipe and add a model to it.

# Set the workflow
rf_trend_workflow <- workflow() |>
  add_recipe(trend_recipe) |>
  add_model(rf_trend_model)

rf_trend_workflow

# Tune hyper parameters by checking model performance using cross validation

# Default mtry is the (rounded down) square root of the number variables and tree is 500. Specify which values want to try in a grid
rf_trend_grid <- expand.grid(
  mtry = seq(2, 5, by = 1),
  trees = seq(200, 1000, by = 200)
  )

# Extract results of tuning (metric is rmse)
rf_trend_tune_results <- rf_trend_workflow |>
  tune_grid(resamples = trend_cv, # CV object
            grid = rf_trend_grid) # grid of values to try

# Print results
rf_trend_tune_results |>
  collect_metrics()

p2 <- autoplot(rf_trend_tune_results) +
  scale_color_brewer(palette = "Set1") + 
  theme(legend.position = "bottom")
# Seems mtry = 3 is the best (just) 

p2 

# Now update the workflow to get the best mtry and n trees from the tuning
param_trend_final <- rf_trend_tune_results |>
  select_best(metric = "rmse")

param_trend_final

# Add this parameter to the workflow using the finalize_workflow() function.
rf_trend_workflow <- rf_trend_workflow |>
  finalize_workflow(param_trend_final) # can do the select best in here as well

rf_trend_workflow

# Now we’ve defined our recipe, our model, and tuned the model’s parameters, we can fit the final model. Applying the `last_fit()` function to our workflow and our train/test split object will train the model specified by the workflow using the training data, and produce evaluations based on the test set.

rf_trend_fit_short <- rf_trend_workflow |>
  last_fit(trend_split)

# Evaluate the model performance on the testing data

rf_trend_fit_short |> collect_metrics()

# Now that we have a final model we want to train it on the full dataset, and then use it to predict the response for new data. For that we need to `fit()` again on the full dataset (the complete training + testing dataset).

final_trend_model <- fit(rf_trend_workflow, trend_short)

final_trend_model

# Save!
saveRDS(final_trend_model, paste0(home, "/output/rf_final_trend_model_short.rds"))
```

Fit plots

```{r}
# Plot results
# FIXME
# I want to combine these to GGSAVE plots into A/B
# ggsave(paste0(home, "/figures/supp/hyper_tuning_old.pdf"), width = 11, height = 11, units = "cm")

# Check predicted vs observed, note this is for the testing data!
p1 <- collect_predictions(rf_trend_fit_old) |>
  ggplot(aes(biom_trend, .pred)) +
  geom_point(alpha = 0.5, color = "steelblue") +
  facet_wrap(~"1978-2023\n(Cod, Flounder, Plaice)") +
  labs(x = "Biomass trend") +
  geom_abline(linetype = 2, color = "tomato3") +
  NULL

p2 <- collect_predictions(rf_trend_fit_old) |>
  ggplot(aes(biom_trend, .pred)) +
  geom_point(alpha = 0.5, color = "steelblue") +
  facet_wrap(~"1993-2023\n(Whiting, Turbot, Dab)") +
  labs(x = "Biomass trend") +
  geom_abline(linetype = 2, color = "tomato3") +
  NULL

(p1 + p2) + plot_layout(guides = "collect", axes = "collect")

ggsave(paste0(home, "/figures/supp/obs_pred.pdf"), width = 18, height = 10, units = "cm")
```

Load model

```{r}
final_trend_model_old <- readRDS(paste0(home, "/output/rf_final_trend_model_old.rds"))
final_trend_model_short <- readRDS(paste0(home, "/output/rf_final_trend_model_short.rds"))
```

Look at variable importance

```{r trend vip}
#| message: false
# p <- 
#   # TODO: or trend_train?
#   final_trend_model_old |> 
#   extract_fit_parsnip() |>
#   vip(aesthetics = list(alpha = 0.8, fill = "steelblue"))

# For a tailor made plot we can also extract the VIPs
trend_ranger_obj_old <- extract_fit_parsnip(final_trend_model_old)$fit
trend_ranger_obj_short <- extract_fit_parsnip(final_trend_model_short)$fit
#trend_ranger_obj$variable.importance

create_vip_tibble <- function(trend_ranger_obj) {
  tibble(Importance = trend_ranger_obj$variable.importance,
         var = as.factor(names(trend_ranger_obj$variable.importance))) |> 
    mutate(var2 = recode(var, 
                         "mean_biom" = "Mean biomass density",
                         "oxy_vel" = "Oxygen velocity",
                         "temp_vel" = "Temperature velocity",
                         "mean_temp" = "Mean temperature",
                         "mean_oxy" = "Mean oxygen",
                         "group" = "Group"),
           response = "Trend")
}

vip_tib_trend_old <- create_vip_tibble(trend_ranger_obj_old)
vip_tib_trend_short <- create_vip_tibble(trend_ranger_obj_short)

vip_tib_trend <- bind_rows(
  vip_tib_trend_old |> mutate(model = "1978-2023\n(Cod, Flounder)"),
  vip_tib_trend_short |> mutate(model = "1993-2023\n(Plaice, Whiting, Turbot, Dab)")
)

vip1 <- ggplot(vip_tib_trend |> filter(model == "1978-2023\n(Cod, Flounder)"),
               aes(Importance, reorder(var2, Importance))) + 
  geom_point() + 
  annotate("text", label = "(a)", x = -Inf, y = Inf, vjust = 1.5, hjust = -0.2, color = "gray40", size = 3.5) +
  facet_wrap(~model) + 
  geom_segment(aes(x = 0, xend = Importance,
                   yend = as.numeric(reorder(var2, Importance)))) +
  theme(axis.title.y = element_blank())

vip2 <- ggplot(vip_tib_trend |> filter(model == "1993-2023\n(Plaice, Whiting, Turbot, Dab)"),
               aes(Importance, reorder(var2, Importance))) + 
  annotate("text", label = "(b)", x = -Inf, y = Inf, vjust = 1.5, hjust = -0.2, color = "gray40", size = 3.5) +
  geom_point() + 
  facet_wrap(~model) + 
  geom_segment(aes(x = 0, xend = Importance,
                   yend = as.numeric(reorder(var2, Importance)))) +
  theme(axis.title.y = element_blank())


## Plot conditional effects and partial dependence plots

# Conditional effect of temperature (using the range in the old data)
nd_temp_old <- trend_old |>
  group_by(group) |> 
  data_grid(mean_temp = seq_range(mean_temp, n = 100, trim = 0.1)) |> 
  ungroup() |> 
  mutate(temp_vel = mean(d_old$temp_vel),
         oxy_vel = mean(d_old$oxy_vel),
         mean_biom = mean(d_old$mean_biom),
         mean_oxy = mean(d_old$mean_oxy),
         group2 = str_replace(group, "_", " "),
         group2 = str_to_sentence(group2)) |> 
  separate_wider_delim(group2, delim = " ", names = c("Species", "Life stage"), 
                       cols_remove = FALSE) |> 
  mutate(`Life stage` = str_to_sentence(`Life stage`))

# Predict
nd_temp_old$pred <- predict(final_trend_model_old, new_data = nd_temp_old)$.pred

# Predict and get jacknife CI's
# So, we don't have a ranger object but a workflow object... 
# https://workflows.tidymodels.org/reference/predict-workflow.html
# I assume because this is a ranger model, the se.method = "infjack"
# https://github.com/imbs-hl/ranger/issues/136
pred_ci_old <- predict(final_trend_model_old, new_data = nd_temp_old, type = "conf_int", level = 0.95)

nd_temp_old <- nd_temp_old |> bind_cols(pred_ci_old)
  

# Now the shorter time frame
nd_temp_short <- trend_short |>
  group_by(group) |> 
  data_grid(mean_temp = seq_range(mean_temp, n = 100, trim = 0.1)) |> 
  ungroup() |> 
  mutate(temp_vel = mean(d_short$temp_vel),
         oxy_vel = mean(d_short$oxy_vel),
         mean_biom = mean(d_short$mean_biom),
         mean_oxy = mean(d_short$mean_oxy),
         group2 = str_replace(group, "_", " "),
         group2 = str_to_sentence(group2)) |> 
  separate_wider_delim(group2, delim = " ", names = c("Species", "Life stage"), 
                       cols_remove = FALSE) |> 
  mutate(`Life stage` = str_to_sentence(`Life stage`))
  
nd_temp_short$pred <- predict(final_trend_model_short, new_data = nd_temp_short)$.pred

pred_ci_short <- predict(final_trend_model_short, new_data = nd_temp_short, type = "conf_int", level = 0.95)

nd_temp_short <- nd_temp_short |> bind_cols(pred_ci_short)

nd_temp <- bind_rows(nd_temp_old, nd_temp_short) |> 
  mutate(pred_ct = pred - mean(pred),
         .pred_lower_ct = .pred_lower - mean(.pred_lower),
         .pred_upper_ct = .pred_upper - mean(.pred_upper),
         .by = group)

hist(nd_temp$pred)

pal <- brewer.pal(n = 12, name = "Paired")[c(1:6, 8, 10, 12)]

# Plot conditional effect of temperature
cond_temp <- ggplot(nd_temp, aes(mean_temp, pred_ct, color = group2)) + 
  geom_ribbon(aes(ymin = .pred_lower_ct, ymax = .pred_upper_ct, fill = group2),
              alpha = 0.15, color = NA) +
  geom_line(alpha = 0.8, linewidth = 0.9) + 
  facet_wrap(~"Conditional effect") +
  scale_color_manual(values = pal) +
  scale_fill_manual(values = pal) +
  labs(x = "Mean temperature (°C)", 
       y = "Scaled local biomass trend") + 
  geom_smooth(aes(mean_temp, pred_ct), inherit.aes = FALSE,
              color = "black", alpha = 0.4, size = 0.75) +
  annotate("text", label = "(c)", x = -Inf, y = Inf, vjust = 1.5, hjust = -0.2, color = "gray40", size = 4) +
  guides(fill = "none",
          color = "none")

cond_temp

# Partial dependence plot
# https://stackoverflow.com/questions/67634344/r-partial-dependence-plots-from-workflow
explainer_rf_old <- explain_tidymodels(
    final_trend_model_old, 
    data = dplyr::select(trend_old, -biom_trend), 
    y = trend_old$biom_trend,
    label = "random forest"
)

pdp_rf_old <- model_profile(explainer_rf_old, N = NULL, 
                            variables = c("mean_temp", "mean_oxy", "temp_vel",
                                          "oxy_vel", "mean_biom"),
                            groups = "group")

explainer_rf_short <- explain_tidymodels(
    final_trend_model_short, 
    data = dplyr::select(trend_short, -biom_trend), 
    y = trend_short$biom_trend,
    label = "random forest"
)

pdp_rf_short <- model_profile(explainer_rf_short, N = NULL, 
                              variables = c("mean_temp", "mean_oxy", "temp_vel",
                                            "oxy_vel", "mean_biom"),
                              groups = "group")

pd <- bind_rows(
  as_tibble(pdp_rf_old$agr_profiles),
  as_tibble(pdp_rf_short$agr_profiles)) |> 
  janitor::clean_names() |> 
  mutate(var2 = recode(vname,
                       "Mean biom" = "Mean biomass density",
                       "Oxy vel" = "Oxygen velocity",
                       "Temp vel" = "Temperature velocity",
                       "Mean temp" = "Mean temperature",
                       "Mean oxy" = "Mean oxygen",
                       "Group" = "Group"))

# Trime the range of x-values
pd <- pd |>
  mutate(group = str_to_sentence(str_replace(groups, "_", " ")),
         vname = str_to_sentence(str_replace(vname, "_", " "))) |> 
  mutate(lwr = quantile(x, probs = 0.05),
         upr = quantile(x, probs = 0.95),
         .by = c(group, vname)) |> 
  filter(x > lwr & x < upr) |> 
  mutate(yhat_ct = yhat - mean(yhat), .by = group)

# How the value is changing with the change in the given predictor variable 
pal <- brewer.pal(n = 12, name = "Paired")[c(1:6, 8, 10, 12)]

part_dep <- pd |> 
  filter(vname == "Mean temp") |> 
  ggplot(aes(x, yhat_ct, color = group)) +
  geom_line(alpha = 0.8, linewidth = 0.9) +
  scale_color_manual(values = pal) +
  labs(x = "Mean temperature (°C)", 
       y = "Scaled local biomass trend") + 
  facet_wrap(~"Partial dependence") +
  geom_smooth(aes(x, yhat_ct), inherit.aes = FALSE,
              color = "black", alpha = 0.4, size = 0.75) +
  annotate("text", label = "(d)", x = -Inf, y = Inf, vjust = 1.5, hjust = -0.2, color = "gray40", size = 4) +
  guides(color = guide_legend(hjust = -10))

p <- (cond_temp + part_dep) +
  plot_layout(guides = "collect",
              axes = "collect") & 
  theme(legend.position = "bottom",
        legend.margin = margin(t = 0, r = 0, b = 0, l = 0),
        legend.box.margin = margin(t = -0, r = 0, b = 0, l = 0),
        legend.spacing.x = unit(0, 'cm'),
        legend.spacing.y = unit(0, 'cm'),
        legend.title = element_blank())

vip <- (vip1 | vip2) +
  plot_layout(axes = "collect")

vip / free(p) +
  plot_layout(heights = c(1, 2))

ggsave(paste0(home, "/figures/effects.pdf"), width = 18, height = 17, units = "cm")
```

```{r}
part_dep_all <- pd |> 
  ggplot(aes(x, yhat_ct, color = group)) +
  geom_line(alpha = 0.8) +
  facet_wrap(~vname, ncol = 3, scales = "free") +
  scale_color_manual(values = pal) +
  labs(x = "") +
  geom_smooth(aes(x, yhat_ct), inherit.aes = FALSE,
              color = "black", alpha = 0.4, linewith = 0.75, size = 0.75) +
  theme(legend.position = c(0.85, 0.25),
        legend.title = element_blank())

part_dep_all

ggsave(paste0(home, "/figures/supp/partial_dependence.pdf"), width = 18, height = 11, units = "cm")
```

